### 2024-03-13 study_1/analysis.Rmd
analysis experiment 1

---
title: "Analysis Study 1"
output:
  bookdown::pdf_document2: default
  bookdown::word_document2: default
  bookdown::html_document2:
      keep_md: yes
always_allow_html: true
bibliography: ../references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r packages, message=FALSE, echo=FALSE, include=FALSE}
library(tidyverse)     # create plots with ggplot, manipulate data, etc.
library(broom.mixed)   # convert regression models into nice tables
library(modelsummary)  # combine multiple regression models into a single table
library(lme4)          # model specification / estimation 
library(lmerTest)      # provides p-values in the output
library(ggpubr)        # stile feature of ggplot
library(gghalves)      # do special plots in ggplot
library(kableExtra)    # for tables
library("grid")        # for image plots   
library("gridExtra")   # for image plots
library("png")         # for image plots
library(stargazer)
```

```{r, include=FALSE, message=FALSE}
# read the cleaned long version data set
data_long <- read_csv("data/cleaned_long.csv")

# read wide version data set
data_wide <- read_csv("data/cleaned_wide.csv")

# make a version with averages per participant
d_avg <- data_long %>% 
  # make numeric versions
  mutate(acceptance_num = ifelse(acceptance == "Yes", 1, 0), 
         knowledge_num = ifelse(knowledge == TRUE, 1, 0)
  ) %>% 
  group_by(id) %>% 
  # calculate by-participant averages
  summarize(avg_acceptance  = sum(acceptance_num)/n(), 
            avg_knowledge = sum(knowledge_num)/n()
  )

# make data frame with average particpant data and wide format
d <- left_join(data_wide %>% 
                 select(-c(ends_with("know"), ends_with("accept"))), 
               d_avg, 
               by = "id")

```

# RQ1

**RQ1: What is the average science knowledge score (1)?**

```{r}
# as table
# d_avg %>% 
#   mutate(avg_knowledge = as.factor(avg_knowledge)) %>% 
#   group_by(avg_knowledge) %>% 
#   count() %>% 
#   ungroup() %>% 
#   mutate(share = n / sum(n)) %>% 
#   mutate_if(is.numeric, round, digits = 3)

# as plot
# Calculate the mean
mean_value <- mean(d_avg$avg_knowledge)

ggplot(d_avg, aes(x = avg_knowledge)) +
  geom_bar() + 
  geom_vline(xintercept = mean_value, color = "red", linetype = "dashed") +
  geom_text(aes(x = mean_value, y = 40, label = paste("Mean =", round(mean_value, 2))), 
            vjust = -0.5, hjust = 1.2) +
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  labs(x = "Average science knowledge", 
       y = "N")
```

# RQ2

**RQ2: What is the average acceptance of the scientific consensus (2)?**

```{r}
# Calculate the mean
mean_value <- mean(d_avg$avg_acceptance)

ggplot(d_avg, aes(x = avg_acceptance)) +
  geom_bar() + 
  geom_vline(xintercept = mean_value, color = "red", linetype = "dashed") +
  geom_text(aes(x = mean_value, y = 60, label = paste("Mean =", round(mean_value, 2))), 
            vjust = -0.5, hjust = 1.2) +
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  labs(x = "Average consensus acceptance", 
       y = "N")
```

# RQ3

**RQ3: What is the relationship between trust in science and, respectively, knowledge and acceptance?**

```{r}
plot_data <- d %>% 
  group_by(wgm_sciencegeneral, avg_knowledge) %>% 
  summarise(n = n())

# Plot as a tile plot
ggplot(plot_data, aes(x = avg_knowledge, y = wgm_sciencegeneral, fill = n)) +
  geom_tile() +  # Add black border around tiles for better visibility
  geom_jitter(inherit.aes = FALSE, data = d, aes(x = avg_knowledge, y = wgm_sciencegeneral),
              width = 0.03, height = 0.08) +  # Scatter points within each tile
  scale_fill_viridis_c(option = "plasma") +  # Use Viridis color scale (Plasma)
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  scale_y_continuous(breaks = 1:5) +
  labs(x = "Average science knowledge", 
       y = "Trust in science", 
       fill = "N") +  # Update fill legend title
  theme_minimal() 
```

```{r}
plot_data <- d %>% 
  group_by(wgm_sciencegeneral, avg_acceptance) %>% 
  summarise(n = n())

# Plot as a tile plot
ggplot(plot_data, aes(x = avg_acceptance, y = wgm_sciencegeneral, fill = n)) +
  geom_tile() +  # Add black border around tiles for better visibility
  geom_jitter(inherit.aes = FALSE, data = d, aes(x = avg_acceptance, y = wgm_sciencegeneral),
              width = 0.03, height = 0.08) +  # Scatter points within each tile
  scale_fill_viridis_c(option = "plasma") +  # Use Viridis color scale (Plasma)
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  scale_y_continuous(breaks = 1:5) +
  labs(x = "Average consensus Acceptance", 
       y = "Trust in science", 
       fill = "N") +  # Update fill legend title
  theme_minimal() 
```

# RQ4

**RQ4: What is the relationship between conspiracy thinking and, respectively, knowledge and acceptance?**

```{r}
# Plot as a tile plot
ggplot(d, aes(x = avg_knowledge, y = BCTI_avg)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  labs(x = "Average science knowledge", 
       y = "Average conspiracy thinking") +  # Update fill legend title
  theme_minimal() 
```

```{r}
ggplot(d, aes(x = avg_acceptance, y = BCTI_avg)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  labs(x = "Average consensus acceptance", 
       y = "Average conspiracy thinking") +  # Update fill legend title
  theme_minimal() 
```

## Difference between false and true answer acceptance

```{r}
# make a version with averages per participant, grouped by false/true ansers 
false_data_avg <- data_long %>% 
  # make numeric versions
  mutate(acceptance_num = ifelse(acceptance == "Yes", 1, 0)
  ) %>% 
  group_by(id, knowledge) %>% 
  # calculate by-participant averages
  summarize(
    n = n(),
    n_accepted = sum(acceptance_num),
    avg_acceptance  = sum(acceptance_num)/n(), 
    BCTI_avg = mean(BCTI_avg)
  ) %>% 
  mutate(knowledge = ifelse(knowledge == FALSE, "False answers", "Correct answers"))

# make data frame with average particpant data and wide format
# false_data <- left_join(false_data, false_data_avg, by = "id")

ggplot(false_data_avg, aes(x = avg_acceptance, y = BCTI_avg)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  labs(x = "Average consensus acceptance", 
       y = "Average conspiracy thinking") +  # Update fill legend title
  theme_minimal() +
  facet_wrap(~knowledge)

```

Alternative plot

```{r}
plot_data <- data_long %>% 
  mutate(BCTI_avg = round(BCTI_avg, digits = 0)) %>% 
  group_by(knowledge, acceptance, BCTI_avg) %>% 
  count() %>% 
  group_by(knowledge, BCTI_avg) %>% 
  mutate (share = n/sum(n)) %>% 
  mutate_if(is.numeric, round, digits = 3) %>% 
  mutate(knowledge = ifelse(knowledge == FALSE, "False answers", "Correct answers"))

ggplot(plot_data, aes(x = BCTI_avg, y = share, fill = acceptance)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
  scale_x_continuous(breaks = seq(1, 9, 1)) +
  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
            position = position_dodge(width = 0.9),
            vjust = -0.5, size = 3) +
  scale_fill_viridis_d(option = "cividis") +
  labs(x = "Average conspiracy thinking", fill = "Acceptance \nof consensus") +
  theme_minimal() +
  facet_wrap(~knowledge) +
  theme(legend.position = "top")

ggplot(data_long, aes(x = ))
```

# Additional RQs

## Knowledge and Acceptance by science question

```{r}
# calculate share
plot_data <- data_long %>% 
  group_by(subject) %>% 
  summarize(share_true = sum(knowledge == TRUE, na.rm = TRUE)/n(),
            share_accepted = sum(acceptance == "Yes", na.rm = TRUE)/n()) 
```

```{r}
# plot knowledge
ggplot(plot_data, aes(x = reorder(subject, share_true), y = share_true)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
  labs(x = "Question", y = "Share of correct responses") +
  coord_flip()
```

```{r}
# plot acceptance
ggplot(plot_data, aes(x = reorder(subject, share_accepted), y = share_accepted)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
  labs(x = "Question", y = "Share of acceptance") +
  coord_flip()
```

## Open ended comments
```{r}
levels(as.factor(d$comments))
```

## Is the share of correct responses associated with the share of acceptance ?

```{r}
plot_data <- d %>% 
  group_by(avg_knowledge, avg_acceptance) %>% 
  summarise(n = n())

# Plot as a tile plot
ggplot(plot_data, aes(x = avg_knowledge, y = avg_acceptance, fill = n)) +
  geom_tile() +  # Add black border around tiles for better visibility
  geom_jitter(inherit.aes = FALSE, data = d, aes(x = avg_knowledge, y = avg_acceptance),
              width = 0.03, height = 0.04) +  # Scatter points within each tile
  scale_fill_viridis_c(option = "plasma") +  # Use Viridis color scale (Plasma)
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  scale_y_continuous(breaks = seq(0, 1, 0.1)) +
  labs(x = "Average science knowledge", 
       y = "Average consensus acceptance", 
       fill = "N") +  # Update fill legend title
  theme_minimal() 
```

# What were the acceptance rates among true and false answers?

```{r}
plot_data <- data_long %>% 
  group_by(knowledge, acceptance) %>% 
  count() %>% 
  group_by(knowledge) %>% 
  mutate (share = n/sum(n)) %>% 
  mutate_if(is.numeric, round, digits = 3)

ggplot(plot_data, aes(x = knowledge, y = share, fill = acceptance)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
            position = position_dodge(width = 0.9),
            vjust = -0.5, size = 3) +
  scale_fill_viridis_d(option = "cividis") +
  labs(x = "Knowledge answer", fill = "Acceptance \nof consensus") +
  theme_minimal()
```

## To which extent did people believe the different conspiracy items ?

```{r}
BCTI_data <- data_long %>% 
  # bring to long format
  pivot_longer(cols = BCTI_apollo:BCTI_evolution, 
               names_to = "BCTI_item", values_to = "score") %>% 
  select(id, c(starts_with("BCTI"), score)) %>% 
  group_by(BCTI_item) %>% 
  summarize(avg_score = mean(score)) %>% 
  ungroup()

ggplot(BCTI_data, aes(x = reorder(BCTI_item, avg_score), y = avg_score)) +
    geom_pointrange(aes(ymin = 1, ymax = avg_score),
                  # Make the lines a little thicker and the dots a little bigger
                  fatten = 2, size = 1) +
  scale_y_continuous(limits = c(1, 9), breaks = seq(1, 9, 1)) +
  labs(x = "Conspiracy item", y = "Average score") +
  coord_cartesian(ylim = c(1, 9)) +
  coord_flip()
```