### 2024-03-25 appendix_exp1.Rmd
draft after study 1

# Experiment 1 {#exp1}

## Materials

\FloatBarrier

### Conspiracy scales

Beside Belief in Conspiracy Theory Inventory (BCTI) by @pennycookOverconfidentlyConspiratorialConspiracy2022 which report in the main study, we also assessed two other consipiracy thinking measures:

1. The conspiracy mentality questionnaire (CMQ) by @bruderMeasuringIndividualDifferences2013:
I think that . . .

-   ... many very important things happen in the world, which the public is never informed about. - politicians usually do not tell us the true motives for their decisions.
-   ... government agencies closely monitor all citizens.
-   ... events which superficially seem to lack a connection are often the result of secret activities.
-   ... there are secret organizations that greatly influence political decisions.

[0% - 100%; 0 = certainly not, 100 = certain]

2.  The Single Item Conspiracy Beliefs Scale (SICBS) by @lantianMeasuringBeliefConspiracy2016 :

-   I think that the official version of the events given by the authorities very often hides the truth. [1-9; 1 = Completely false, 5 = Unsure, 9 = Completely true]

### Trust in science

We rely on three items

1. How much do you trust scientists in this country? Do you trust them a lot, some, not much, or not at all? [1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]

2. In general, would you say that you trust science a lot, some, not much, or not at all? [1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]

3. How much confidence do you have in scientists to act in the best interests of the public? [1-5; 1 = No confidence at all, 5 = A great deal of confidence]

## Comparing items

Make correlation table

### Conspiracy theories

Table \@ref(tab:correlation-conspiracy) shows the correlations of the three different scales assessing conspiracy thinking.

```{r correlation-conspiracy}
correlation_matrix <- exp1_wide %>%
  select(BCTI_avg, CMQ_avg, SICBS) %>% 
  rename(BCTI = BCTI_avg, 
         CMQ = CMQ_avg) %>% 
  cor()

apa_table(correlation_matrix, 
          caption = "Correlations of the three different scales assessing conspiracy thinking",
          placement = "h")
```

### Trust in science

Table \@ref(tab:correlation-trust) shows the correlations of the three different items measuring trust in science.

```{r correlation-trust}
correlation_matrix <- exp1_wide %>%
  select(wgm_sciencegeneral, wgm_scientists, pew) %>% 
  cor()

apa_table(correlation_matrix, 
          caption = "Correlations of the three different items measuring trust in science",
          placement = "h")
```

## Correlations with alternative measures

Table \@ref(tab:correlations-outcomes) shows the correlations between knowledge and acceptance, respectively, and outcome variables.

```{r}
# Standardize all variables. Linear regression on standardized variables is equivalent to correlation
standardized_variables <- exp1_wide %>% 
  select(avg_acceptance, avg_knowledge, BCTI_avg, CMQ_avg, SICBS, wgm_scientists, wgm_sciencegeneral, pew) %>% 
  mutate(across(everything(), ~scale(.x)))
```

```{r}
# run models to get correlation coefficient

# Define the function to run regression and tidy results
run_regression <- function(data, dependent_var, independent_var) {
  lm(as.formula(paste(dependent_var, "~", independent_var)), data = data) %>%
    tidy() %>%
    filter(term == independent_var) %>%
    mutate(outcome = {{dependent_var}}) %>% 
    select(outcome, term, estimate)
}

# List of variables
variables <- c("BCTI_avg", "CMQ_avg", "SICBS", "wgm_scientists", "wgm_sciencegeneral", "pew")

# Run regression for each variable and store results in a list
results_list_knowledge <- map(variables, ~ run_regression(standardized_variables, .x, "avg_knowledge"))
results_list_acceptance <- map(variables, ~ run_regression(standardized_variables, .x, "avg_acceptance"))
```

```{r correlations-outcomes}
# as data frame/output table
results <- bind_rows(list(results_list_knowledge, results_list_acceptance))
results %>% 
  pivot_wider(names_from = term, 
              values_from = estimate) %>% 
  rename(`Correlation with knowledge` = avg_knowledge, 
         `Correlation with acceptance` = avg_acceptance, 
         ) %>% 
  mutate_if(is.numeric, round, digits = 2) %>% 
  mutate(outcome = case_when(outcome == "BCTI_avg" ~ "BCTI \n(main conspiracy measure)", 
                             outcome == "CMQ_avg" ~ "CMQ", 
                             outcome == "SICBS" ~ "SICBS", 
                             outcome == "wgm_scientists" ~ "WGM trust scientists",
                             outcome == "wgm_sciencegeneral" ~ "WGM trust general \n(main trust measure)",
                             outcome == "pew" ~ "PEW trust scientists",
                             )) %>% 
  apa_table(caption = "Correlations between knowledge and acceptance, respectively, and outcome variables")
```

## Results conditional on false responses

Table \@ref(tab:false-response-regression) shows the correlations between acceptance and outcome variables based on linear regression models on standardized values.

```{r}
# Standardize all variables. Linear regression on standardized variables is equivalent to correlation
standardized_variables <- exp1_false_knowledge %>% 
  select(avg_acceptance, BCTI_avg, CMQ_avg, SICBS, wgm_scientists, wgm_sciencegeneral, pew) %>% 
  mutate(across(everything(), ~scale(.x)))
```

```{r}
# Define the function to run regression and tidy results
run_regression <- function(data, dependent_var, independent_var) {
  lm(as.formula(paste(dependent_var, "~", independent_var)), data = data) 
}

# run models to get correlation coefficient

# List of variables
variables <- c("BCTI_avg", "CMQ_avg", "SICBS", "wgm_scientists", "wgm_sciencegeneral", "pew")

# Run regression for each variable and store results in a list
results_list <- setNames(map(variables, ~ run_regression(standardized_variables, .x, "avg_acceptance")), variables)
```

```{r false-response-regression}
modelsummary(results_list, 
             stars = TRUE,
             caption = "Based on false response data only, correlations between acceptance and outcome variables based on linear regression models on standardized values") 
```