### 2024-03-25 exp_1/cleaning.Rmd

---
title: "Data cleaning Study 1"
output: 
  html_document: 
    keep_md: yes
---

```{r packages, message=FALSE}
library(tidyverse)     # create plots with ggplot, manipulate data, etc.
library(broom.mixed)   # convert regression models into nice tables
library(modelsummary)  # combine multiple regression models into a single table
library(lme4)          # model specification / estimation 
library(lmerTest)      # provides p-values in the output
library(ggpubr)        # stile feature of ggplot
library(gghalves)      # do special plots in ggplot
library(kableExtra)    # for tables
```

## Import data
```{r import, message=FALSE}
d <- read_csv("./data/qualtrics.csv")
names(d)
```

```{r}
# inspect
head(d) # you can also use View(d)
# delete first two rows
d <- d %>% 
  slice(3: nrow(.)) 
```

## Attention check
```{r attention}
# attention check
table(d$attention_check)
```

There are 6 failed attention checks.

```{r}
# filter to only valid attention check responses
d <- d %>% filter(attention_check == "Never 1")
```

## Add demographics

## Recode demographics

```{r, message=FALSE}
prolific_demographics <- read_csv("./data/prolific_demographics.csv")

d <- left_join(d, prolific_demographics, by = c("PROLIFIC_PID" = "Participant id"))
```

```{r}
d <- d %>% 
  mutate(gender = case_when(Sex == "Male" ~ "male", 
                            Sex == "Female" ~  "female", 
                            .default = NA)
         ) %>% 
  rename(age = Age)
```

## Clean and re-shape data

```{r}
# check all names and their order
names(d)
```

Clean wide format data.

```{r}
# Function to extract numbers from a string for BCTI and SICBS
extract_numbers <- function(string_var) {
  as.numeric(gsub("\\D", "", string_var))
}

# test the function
# d %>% 
#   mutate(across(starts_with("BCTI"), ~extract_numbers(.), .names = "{.col}_num")) %>% 
#   select(BCTI_autism, BCTI_autism_num)

# clean and re-shape
d_wide <- d %>% 
  # add an easy to read participant identifier
  mutate(id = 1:nrow(.)) %>%
  # make some character variables numeric
  mutate(across(c(starts_with("BCTI"), starts_with("SICBS"), starts_with("wgm"), 
                  starts_with("pew")), ~extract_numbers(.)
  )
  ) %>% 
  mutate(across(starts_with("CMQ"), as.numeric)
  ) %>% 
  # add average conspiracy measures 
  rowwise() %>% 
  mutate(BCTI_avg = mean(c_across(starts_with("BCTI")
                                  ),
                         na.rm = TRUE
                         ),
         CMQ_avg = mean(c_across(starts_with("CMQ")
                                  ),
                         na.rm = TRUE
                         )
         ) %>% 
  ungroup() 
```

Make long format data.

```{r}
d_long <- d_wide %>% 
  # bring to long format
  pivot_longer(cols = electrons_know:water_accept, 
               names_to = "subject_question", values_to = "correct_accept") %>% 
  # separate conditions into CONVERGENCE_OUTCOME_STIMULUS
  separate_wider_delim(subject_question, "_", names = c("subject", "question")
                       ) %>%
  # necessary because each participant only saw ten out of 11 questions
  drop_na(correct_accept) %>% 
  pivot_wider(names_from = question, values_from = correct_accept) %>% 
# create better variable names
  rename(knowledge = know, 
         acceptance = accept) 
```

Add knowledge and acceptance by-participant averages to wide format data.

```{r}
# make a version with averages per participant
d_avg <- d_long %>% 
  # make numeric versions
  mutate(acceptance_num = ifelse(acceptance == "Yes", 1, 0), 
         knowledge_num = ifelse(knowledge == TRUE, 1, 0)
  ) %>% 
  group_by(id) %>% 
  # calculate by-participant averages
  summarize(avg_acceptance  = sum(acceptance_num)/n(), 
            avg_knowledge = sum(knowledge_num)/n()
  )

# make data frame with average particpant data and wide format
d_wide <- left_join(d_wide %>% 
                 select(-c(ends_with("know"), ends_with("accept"))), 
               d_avg, 
               by = "id")
```

## Export data

```{r}
# wide format
write_csv(d_wide, "data/cleaned_wide.csv")

# long format
write_csv(d_long, "data/cleaned_long.csv")
```