### 2024-05-31 exp_3/analysis.Rmd

---
title: "Analysis Study 3"
output:
  bookdown::pdf_document2: default
  bookdown::word_document2: default
  bookdown::html_document2:
      keep_md: yes
always_allow_html: true
bibliography: ../references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r packages, message=FALSE, echo=FALSE, include=FALSE}
library(tidyverse)     # create plots with ggplot, manipulate data, etc.
library(broom.mixed)   # convert regression models into nice tables
library(modelsummary)  # combine multiple regression models into a single table
library(lme4)          # model specification / estimation 
library(lmerTest)      # provides p-values in the output
library(ggpubr)        # stile feature of ggplot
library(gghalves)      # do special plots in ggplot
library(kableExtra)    # for tables
library("grid")        # for image plots   
library("gridExtra")   # for image plots
library("png")         # for image plots
library(stargazer)
```

```{r, include=FALSE, message=FALSE}
# read the cleaned long version data set
data_long <- read_csv("data/cleaned_long.csv")

# read wide version data set
data_wide <- read_csv("data/cleaned_wide.csv") 
```

```{r}
plot_data <- data_long %>% 
  group_by(knowledge, acceptance) %>% 
  count() %>% 
  group_by(knowledge) %>% 
  mutate (share = n/sum(n)) %>% 
  mutate_if(is.numeric, round, digits = 3)

absolute_numbers <- plot_data %>% 
  group_by(knowledge) %>% 
  summarise(n = sum(n))

ggplot(plot_data, aes(x = knowledge, y = share, fill = acceptance)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
            position = position_dodge(width = 0.9),
            vjust = -0.5, size = 3) +
  scale_fill_viridis_d(option = "cividis") +
  labs(x = "Knowledge answer", fill = "Acceptance \nof consensus") +
  theme_minimal() +
  geom_bracket(data = absolute_numbers,
               aes(xmin = as.numeric(factor(knowledge)) - 0.4, 
                   xmax = as.numeric(factor(knowledge)) + 0.4, 
                   y.position = 1.05, label =  paste0("n = ", n)),
               inherit.aes = FALSE,
               step.increase = 0)
```