### 2024-06-24 appendix_exp1.Rmd
making summary plot and adding some first introduction bits

# Experiment 1 {#exp1}

## Materials

\FloatBarrier

(ref:stimulus-example) Example of a science question, the scientific consensus and the corresponding acceptance question.

```{r stimulus-example, echo=FALSE, out.width= "100%", fig.align="left", fig.show="hold", fig.cap="(ref:stimulus-example)"}
knitr::include_graphics("./figures/study1_question_example.png")
```

```{r questions-with-sources}
# items <- read_csv("exp_1/materials/knowledge_items.csv") %>% 
#   mutate(id = 1:nrow(.)) %>% 
#   select(id, Question, `Scientific consensus`, `Reference(s)`)
# 
# # Output the table
# kbl(items, booktabs = T, longtable = TRUE,
#     caption = "Science knowledge items", 
#     full_width = T) %>%
#   column_spec(1, width = "1em") %>%
#   column_spec(2, width = "10em") %>%
#   column_spec(3, width = "10em") %>%
#   column_spec(4, width = "15em") 
```

### Conspiracy scales

Beside Belief in Conspiracy Theory Inventory (BCTI) by @pennycookOverconfidentlyConspiratorialConspiracy2022 which report in the main study, we also assessed two other consipiracy thinking measures:

1. The conspiracy mentality questionnaire (CMQ) by @bruderMeasuringIndividualDifferences2013:
I think that . . .

-   ... many very important things happen in the world, which the public is never informed about. - politicians usually do not tell us the true motives for their decisions.
-   ... government agencies closely monitor all citizens.
-   ... events which superficially seem to lack a connection are often the result of secret activities.
-   ... there are secret organizations that greatly influence political decisions.

[0% - 100%; 0 = certainly not, 100 = certain]

2.  The Single Item Conspiracy Beliefs Scale (SICBS) by @lantianMeasuringBeliefConspiracy2016 :

-   I think that the official version of the events given by the authorities very often hides the truth. [1-9; 1 = Completely false, 5 = Unsure, 9 = Completely true]

### Trust in science

We rely on three items

1. How much do you trust scientists in this country? Do you trust them a lot, some, not much, or not at all? [1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]

2. In general, would you say that you trust science a lot, some, not much, or not at all? [1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]

3. How much confidence do you have in scientists to act in the best interests of the public? [1-5; 1 = No confidence at all, 5 = A great deal of confidence]

## Comparing items

### Conspiracy theories

Table \@ref(tab:correlation-conspiracy) shows the correlations of the three different scales assessing conspiracy thinking.

```{r correlation-conspiracy}
correlation_matrix <- exp1_wide %>%
  select(BCTI_avg, CMQ_avg, SICBS) %>% 
  rename(BCTI = BCTI_avg, 
         CMQ = CMQ_avg) %>% 
  cor()

apa_table(correlation_matrix, 
          caption = "Correlations of the three different scales assessing conspiracy thinking",
          placement = "h")
```

### Trust in science

Table \@ref(tab:correlation-trust) shows the correlations of the three different items measuring trust in science.

```{r correlation-trust}
correlation_matrix <- exp1_wide %>%
  select(wgm_sciencegeneral, wgm_scientists, pew) %>% 
  cor()

apa_table(correlation_matrix, 
          caption = "Correlations of the three different items measuring trust in science",
          placement = "h")
```

## Correlations with alternative measures

Table \@ref(tab:correlations-outcomes) shows the correlations between knowledge and acceptance, respectively, and outcome variables.

```{r}
# Standardize all variables. Linear regression on standardized variables is equivalent to correlation
standardized_variables <- exp1_wide %>% 
  select(avg_acceptance, avg_knowledge, BCTI_avg, CMQ_avg, SICBS, wgm_scientists, wgm_sciencegeneral, pew) %>% 
  mutate(across(everything(), ~scale(.x)))
```

```{r}
# List of variables
variables <- c("BCTI_avg", "CMQ_avg", "SICBS", "wgm_scientists", "wgm_sciencegeneral", "pew")

# Run regression for each variable and store results in a list
results_list_knowledge <- map(variables, ~ run_regression(standardized_variables, .x, "avg_knowledge"))
results_list_acceptance <- map(variables, ~ run_regression(standardized_variables, .x, "avg_acceptance"))
```

```{r correlations-outcomes}
# as data frame/output table
results <- bind_rows(list(results_list_knowledge, results_list_acceptance))
results %>% 
  filter(term != "intercept") %>%
    mutate_if(is.numeric, round, digits = 2) %>% 
  # make a combined variable of estimate and p.value
  mutate(estimate = paste0(estimate, " (p", p.value, ")")) %>% 
  select(outcome, term, estimate) %>% 
  pivot_wider(names_from = term, 
              values_from = estimate) %>% 
  rename(`Correlation with knowledge` = avg_knowledge, 
         `Correlation with acceptance` = avg_acceptance, 
         ) %>% 
  mutate(outcome = case_when(outcome == "BCTI_avg" ~ "BCTI \n(main conspiracy measure)", 
                             outcome == "CMQ_avg" ~ "CMQ", 
                             outcome == "SICBS" ~ "SICBS", 
                             outcome == "wgm_scientists" ~ "WGM trust scientists",
                             outcome == "wgm_sciencegeneral" ~ "WGM trust general \n(main trust measure)",
                             outcome == "pew" ~ "PEW trust scientists",
                             )) %>% 
  apa_table(caption = "Correlations between knowledge and acceptance, respectively, and outcome variables")
```

## Results conditional on false responses

Table \@ref(tab:false-response-regression) shows the correlations between acceptance and outcome variables based on linear regression models on standardized values.

```{r}
# Standardize all variables. Linear regression on standardized variables is equivalent to correlation
standardized_variables <- exp1_false_knowledge %>% 
  select(avg_acceptance, BCTI_avg, CMQ_avg, SICBS, wgm_scientists, wgm_sciencegeneral, pew) %>% 
  mutate(across(everything(), ~scale(.x)))
```

```{r}
# Run regression for each variable and store results in a list
results_list <- setNames(map(variables, ~ run_regression(standardized_variables, .x, "avg_acceptance", return = "model")), variables)
```

```{r false-response-regression}
modelsummary(results_list, 
             stars = TRUE,
             caption = "Based on false response data only, correlations between acceptance and outcome variables based on linear regression models on standardized values") 
```

## By-question variation

```{r}
# calculate share
plot_data <- exp1_long %>% 
  group_by(subject) %>% 
  summarize(share_true = sum(knowledge == TRUE, na.rm = TRUE)/n(),
            share_accepted = sum(acceptance == "Yes", na.rm = TRUE)/n()) 
```

### Knowledge

(ref:exp1-questions-knowledge) Distribution of correct answers by question.

```{r exp1-questions-knowledge, fig.cap="(ref:exp1-questions-knowledge)"}
# plot knowledge
ggplot(plot_data, aes(x = reorder(subject, share_true), y = share_true)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
  labs(x = "Question", y = "Share of correct responses") +
  coord_flip()
```

### Acceptance

(ref:exp1-questions-acceptance) Distribution of consensus acceptance by question.

```{r exp1-questions-acceptance, fig.cap="(ref:exp1-questions-acceptance)"}
# plot acceptance
ggplot(plot_data, aes(x = reorder(subject, share_accepted), y = share_accepted)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
  labs(x = "Question", y = "Share of acceptance") +
  coord_flip()
```

## Trust in science

(ref:exp1-trust-scientists) Distribution of trust in scientists.

```{r exp1-trust-scientists, fig.cap="(ref:exp1-trust-scientists)"}
# calculate share
plot_data <- exp1_wide %>% 
  group_by(wgm_scientists) %>% 
  summarize(n = n()) %>% 
  ungroup() %>% 
  mutate(share = n/sum(n)) %>% 
  rounded_numbers()

# plot knowledge
ggplot(plot_data, aes(x = wgm_scientists, y = share)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
  labs(x = "Trust in science", y = "Share of participants") +
  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
            position = position_dodge(width = 0.9),
            vjust = -0.5, size = 3) 
```

## Conspiracy thinking

### Distribution

(ref:exp1-conspiracy-distribution) Distribution of conspiracy thinking.

```{r exp1-conspiracy-distribution, fig.cap="(ref:exp1-conspiracy-distribution)"}
# plot knowledge
ggplot(exp1_wide, aes(x = BCTI_avg)) +
  geom_histogram() +
  labs(x = "Average conspiracy thinking (BCTI)", y = "N")
```

### By-item variation

(ref:exp1-conspiracy-items) Distribution of conspiracy thinking by item.

```{r exp1-conspiracy-items, fig.cap="(ref:exp1-conspiracy-items)"}
BCTI_data <- exp1_long %>% 
  # bring to long format
  pivot_longer(cols = BCTI_apollo:BCTI_evolution, 
               names_to = "BCTI_item", values_to = "score") %>% 
  select(id, c(starts_with("BCTI"), score)) %>% 
  group_by(BCTI_item) %>% 
  summarize(avg_score = mean(score)) %>% 
  ungroup()

ggplot(BCTI_data, aes(x = reorder(BCTI_item, avg_score), y = avg_score)) +
    geom_pointrange(aes(ymin = 1, ymax = avg_score),
                  # Make the lines a little thicker and the dots a little bigger
                  fatten = 2, size = 1) +
  scale_y_continuous(limits = c(1, 9), breaks = seq(1, 9, 1)) +
  labs(x = "Conspiracy item", y = "Average score") +
  coord_cartesian(ylim = c(1, 9)) +
  coord_flip()
```

## Open ended comments
```{r}
# levels(as.factor(exp1_wide$comments))
table <- exp1_wide %>% 
  drop_na(comments) %>% 
  select(id, comments)

kbl(table, booktabs = T, longtable = TRUE,
    caption = "Reasons for consensus rejection") %>%
  kable_paper(full_width = F) %>% 
  column_spec(1) %>%
  column_spec(2, width = "40em") 
```

## Additional plots

(ref:exp1-conditional-acceptance) Acceptance rates of scientific consensus, based on whether the initial response to the knowledge question was false or true.

```{r exp1-conditional-acceptance, fig.cap="(ref:exp1-conditional-acceptance)"}
plot_data <- exp1_long %>% 
  group_by(knowledge, acceptance) %>% 
  count() %>% 
  group_by(knowledge) %>% 
  mutate (share = n/sum(n)) %>% 
  mutate_if(is.numeric, round, digits = 3)

absolute_numbers <- plot_data %>% 
  group_by(knowledge) %>% 
  summarise(n = sum(n))

ggplot(plot_data, aes(x = knowledge, y = share, fill = acceptance)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
            position = position_dodge(width = 0.9),
            vjust = -0.5, size = 3) +
  scale_fill_viridis_d(option = "cividis") +
  labs(x = "Knowledge answer", fill = "Acceptance \nof consensus") +
  plot_theme +
  geom_bracket(data = absolute_numbers,
               aes(xmin = as.numeric(factor(knowledge)) - 0.4, 
                   xmax = as.numeric(factor(knowledge)) + 0.4, 
                   y.position = 1.05, label =  paste0("n = ", n)),
               inherit.aes = FALSE,
               step.increase = 0)
```

(ref:exp1-plot-overview) **A** Shows the distribution of science knowledge (left) and acceptance of scientific consensus **B** Shows the relationship between trust in science and science knowledge/acceptance of scientific consensus **C** Shows the relationship between conspiracy thinking and science knowledge/acceptance of scientific consensus

```{r exp1-plot-overview, fig.cap="(ref:exp1-plot-overview)", fig.height= 10, fig.width=10}
# Plot distributions of knowledge and acceptance and their mean

mean_value_knowledge <- mean(exp1_wide$avg_knowledge)
max_n <- max(table(exp1_wide$avg_acceptance))

knowledge_mean <- ggplot(exp1_wide, aes(x = avg_knowledge)) +
  geom_bar() + 
  geom_vline(xintercept = mean_value_knowledge, color = "red", linetype = "dashed") +
  geom_text(aes(x = mean_value_knowledge, y = 40, label = paste("Mean =", round(mean_value_knowledge, 2))), 
            vjust = -0.5, hjust = 1.2, check_overlap = TRUE) +
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  scale_y_continuous(breaks = seq(0, max_n, 10)) +
  coord_cartesian(xlim = c(0, 1), 
                  ylim = c(0, max_n)) +
  labs(x = "Average science knowledge", 
       y = "N")

mean_value <- mean(exp1_wide$avg_acceptance)

acceptance_mean <- ggplot(exp1_wide, aes(x = avg_acceptance)) +
  geom_bar() + 
  geom_vline(xintercept = mean_value, color = "red", linetype = "dashed") +
  geom_text(aes(x = mean_value, y = 60, label = paste("Mean =", round(mean_value, 2))), 
            vjust = -0.5, hjust = 1.2, check_overlap = TRUE) +
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  scale_y_continuous(breaks = seq(0, max_n, 10)) +
  coord_cartesian(xlim = c(0, 1), 
                  ylim = c(0, max_n)) +
  labs(x = "Average consensus acceptance", 
       y = "N")

# Trust in science plots
plot_data <- exp1_wide %>% 
  group_by(wgm_sciencegeneral, avg_knowledge) %>% 
  summarise(n = n())

trust_knowledge <- ggplot(plot_data, aes(x = avg_knowledge, y = wgm_sciencegeneral, fill = n)) +
  geom_tile() +  # Add black border around tiles for better visibility
  geom_jitter(inherit.aes = FALSE, data = exp1_wide, aes(x = avg_knowledge, y = wgm_sciencegeneral),
              width = 0.03, height = 0.08) +  # Scatter points within each tile
  scale_fill_viridis_c(option = "plasma") +  # Use Viridis color scale (Plasma)
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  coord_cartesian(xlim = c(0, 1)) +
  scale_y_continuous(breaks = 1:5) +
  labs(x = "Average science knowledge", 
       y = "Trust in science", 
       fill = "N")

plot_data <- exp1_wide %>% 
  group_by(wgm_sciencegeneral, avg_acceptance) %>% 
  summarise(n = n())

trust_acceptance <- ggplot(plot_data, aes(x = avg_acceptance, y = wgm_sciencegeneral, fill = n)) +
  geom_tile() +  
  geom_jitter(inherit.aes = FALSE, data = exp1_wide, aes(x = avg_acceptance, y = wgm_sciencegeneral),
              width = 0.03, height = 0.08) + 
  scale_fill_viridis_c(option = "plasma") + 
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  coord_cartesian(xlim = c(0, 1)) +
  scale_y_continuous(breaks = 1:5) +
  labs(x = "Average consensus Acceptance", 
       y = "Trust in science", 
       fill = "N") +
  theme(legend.position = "top")

# Conspiracy Thinking plots

conspiracy_knowledge <- ggplot(exp1_wide, aes(x = avg_knowledge, y = BCTI_avg)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  scale_y_continuous(breaks = 1:9) +
  coord_cartesian(xlim = c(0, 1), 
                  ylim = c(0, 9)) +
  labs(x = "Average science knowledge", 
       y = "Average conspiracy thinking")

conspiracy_acceptance <- ggplot(exp1_wide, aes(x = avg_acceptance, y = BCTI_avg)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  scale_y_continuous(breaks = 1:9) +
  coord_cartesian(xlim = c(0, 1), 
                  ylim = c(0, 9)) +
  labs(x = "Average consensus acceptance", 
       y = "Average conspiracy thinking") +
  theme(legend.position = "top")

# Combine plots

# distributions
distributions <- ggarrange(knowledge_mean + 
                             rremove("xlab") , 
                           acceptance_mean + 
                             rremove("xlab") +
                             rremove("ylab"))

# trust
trust <- ggarrange(trust_knowledge + 
                             rremove("xlab") +
                             guides(fill = "none") +
                             labs(caption = NULL) , 
                           trust_acceptance + 
                             rremove("xlab") +
                             rremove("ylab"), 
                   common.legend = TRUE)

# conspiracy
conspiracy <- ggarrange(conspiracy_knowledge, 
                           conspiracy_acceptance + 
                             rremove("ylab"), 
                   common.legend = TRUE)

# common plot
patchwork <- (distributions / trust / conspiracy) + 
  plot_annotation(tag_levels = 'A')

patchwork

```