### 2024-09-02 paper.Rmd

---
title             : "Universal acceptance of (nearly all of) basic science in the US"
shorttitle        : "Acceptance of basic science"

header-includes:  | # to prevent floats from moving past certain points (for the appendix)
  \usepackage{placeins}

author: 
  - name          : ""
    affiliation   : ""

affiliation:
  - id            : ""
    institution   : ""

abstract: |
  XX

keywords          :

floatsintext      : yes
linenumbers       : no 
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "doc" # "doc" for nice look, "man" for manuscripty
output            : papaja::apa6_pdf # "_doc" for word; however, note that some of the features of kableExtra are not available for word and will yield errors. For example to knit to word, you'll have to comment out all "add_header_above()" functions for model output tables, or e.g. set always allow html as true in the yaml heading

always_allow_html: true

appendix:
  - "appendix_exp1.Rmd"
  - "appendix_exp2.Rmd"
  - "appendix_exp3.Rmd"
  - "appendix_exp4.Rmd"

bibliography: references.bib
---

```{r setup, include=FALSE}
# Figure out output format
is_docx <- knitr::pandoc_to("docx") | knitr::pandoc_to("odt")
is_latex <- knitr::pandoc_to("latex")
is_html <- knitr::pandoc_to("html")

# Word-specific things
table_format <- ifelse(is_docx, "huxtable", "kableExtra")  # Huxtable tables
conditional_dpi <- ifelse(is_docx, 300, 300)  # Higher DPI
conditional_align <- ifelse(is_docx, "default", "center")  # Word doesn't support align

# Knitr options
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  # tidy.opts = list(width.cutoff = 120),  # Code width
  # fig.retina = 3, dpi = conditional_dpi,
  # fig.width = 7, fig.asp = 0.618,
  # fig.align = conditional_align, out.width = "100%",
  fig.path = "output/figures/",
  cache.path = "output/_cache/",
  fig.process = function(x) {  # Remove "-1" from figure names
    x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
    if (file.rename(x, x2)) x2 else x
  },
  options(scipen = 99999999)  # Prevent scientific notation
)

# R options
options(
  width = 90,  # Output width
  dplyr.summarise.inform = FALSE,  # Turn off dplyr's summarize() auto messages
  knitr.kable.NA = "",  # Make NAs blank in kables
  kableExtra.latex.load_packages = FALSE,  # Don't add LaTeX preamble stuff
  modelsummary_factory_default = table_format,  # Set modelsummary backend
  modelsummary_format_numeric_latex = "plain"  # Don't use siunitx
)
```

```{r packages, include=FALSE}
# load required packages
library("papaja")      # For APA style manuscript   
library("lme4")        # model specification / estimation
library("lmerTest")    # provides p-values in the output
library("tidyverse")   # data wrangling and visualisation
library("afex")        # anova and deriving p-values from lmer
library("broom")       # extracting data from model fits 
library("broom.mixed") # extracting data from mixed models
library("metafor")     # doing mata analysis
library("patchwork")   # put several plots together
library("ggridges")    # for plots
library("gghalves")    # for plots
library("ggbeeswarm")  # Special distribution-shaped point jittering
library("knitr")       # for tables
library("kableExtra")  # also for tables
library("ggpubr")      # for combining plots with ggarrange() 
library("grid")        # for image plots   
library("gridExtra")   # for image plots
library("png")         # for image plots
library("modelsummary") # for regression tables
library("ggrepel") # for better random patterns in plots
```

```{r functions}
# load plot theme
source("functions/plot_theme.R")

# load other functions
source("functions/own_functions.R")
```

# Introduction

How much do people trust and accept science? The recent COVID pandemic has highlighted the importance of this question, with trust in science being the best predictor of people’s proclivity to follow health guidelines [@alganTrustScientistsTimes2021].

Unfortunately, trust in science is far from being at ceiling [@wellcomeglobalmonitorWellcomeGlobalMonitor2018; @wellcomeglobalmonitorWellcomeGlobalMonitor2020], and there are suggestions that it has recently been dropping in some countries [@alganTrustScientistsTimes2021], including the US [@brianAmericansTrustScientists2023], where it is also increasingly politicized [@gauchatPoliticizationSciencePublic2012; @krauseTrendsAmericansTrust2019a; @liPolarizationPublicTrust2022].

This relative lack of trust in science could mean different things. It could mean that some people reject science wholesale, or at least to a significant extent. Or it could mean that they maintain a high degree of trust in most of science, and only question specific results that happen to contradict specific other beliefs they have (e.g. creationism), or beliefs that are associated with disliked behaviors (e.g. vaccination, acting for climate change).

We answer this question by looking at how much Americans accept basic scientific knowledge. We pay special attention to two groups of participants: people who say they don’t trust science (much, or at all), and people who endorse science-related conspiracy theories (CTs), or have a CT mindset, since this has been linked with a low degree of trust in science, and that many CTs imply a mistrust of at least some science (vaccination, hydroxychloroquine, etc.).

We briefly review work on trust in science, and science knowledge, before turning to work on CT and trust in science.

## Trust in science

Across the globe, most people say they trust science at least to some extent. In 2018, the Wellcome Global Monitor (WGM) surveyed of over 140'000 people in over 140 countries on trust in science [@wellcomeglobalmonitorWellcomeGlobalMonitor2018]. In 2020, during the first year of the Covid pandemic, a follow up survey was made in 113 countries, involving 119'000 participants [@wellcomeglobalmonitorWellcomeGlobalMonitor2020]. Across all countries, in 2018, 32% of participants said they trust science "a lot" (41% in 2020), 45% trust "some" science (39% in 2020), and only 13% percent trust science "not much/ not at all" (13% in 2020), while 10% indicated "don't know" (7% in 2020)[^1].

[^1]: <https://wellcome.org/reports/wellcome-global-monitor-covid-19/2020#gid=6acd&pid=0>\>

Trust in science also appears to be relatively stable in the US [@funkPublicConfidenceScientists2020]. However, there are some potentially worrying trends: Recently, trust in US has dropped [@brianAmericansTrustScientists2023], and some evidence suggests increasing polarization of trust in science along political lines [@gauchatPoliticizationSciencePublic2012]. For example, the partisan divide on the question of whether climate change is a major threat to society has been growing in the last decade[^2] [@kennedyWhatDataSays2023]. A recent study suggests that by some measure, only about half of the US population believes in anthropogenic climate change, and that individuals do mostly not seem to change their mind on on the issue [@mottaChangingMindsChanging2021].

[^2]: <https://www.pewresearch.org/short-reads/2023/08/09/what-the-data-says-about-americans-views-of-climate-change/>

Trust in science is typically measured in three ways: either asking explicitly about general trust in science or scientists, by asking about science attitudes (e.g. support for public funding of science), or asking questions about specific, typically contentious, science topics (e.g. vaccines, climate change, GMOs). It is unclear to which extent these measures capture trust in basic science. To assess this, we can look at the literature on science knowledge.

## Science knowledge

Much of the research on public understanding of science measures to which extent people have basic science knowledge (e.g. "Does the Earth go around the Sun, or does the Sun go around the Earth?"). For us, these surveys provide a lower bound to how much people trust basic science: people who give the right answer (and not by chance) supposedly trust science on that question. Unfortunately, that lower bound has often been indeed pretty low: researchers have pointed out the deficit in basic science knowledge [@millerPublicUnderstandingAttitudes2004; @millerMeasurementCivicScientific1998a]. For example, only 55% in the US (in 2014) and only 46% in the EU (in 2005) knew that antibiotics kill only bacteria and not viruses [@committeeonscienceliteracyandpublicperceptionofscienceScienceLiteracyConcepts2016]. A question that remains is whether people do not provide the right answers because they are not familiar with the scientific consensus, or because they explicitly reject that consensus.

[To do: Make a descriptive overview of science knowledge across time/countries.]

## CT and TiS

Recently, the relationship of one group to trust in science (TiS) has been particularly scrutinized: conspiracy theorists (CTists), either defined as people who endorse specific conspiracy theories, or as people who have a conspiracy mindset [@rutjensConspiracyBeliefsScience2022; @vranicDidMyOwn2022]. CTists reject specific science knowledge (related to their CT), and they tend to trust science less on the whole. What remains unclear is to which extent CTists (dis)trust basic science.

## The present studies

In a series of four pre-registered studies (n = 782), we asked participants about basic (studies 1, 2 and 3) and recent/advanced (study 4) science knowledge questions, informed them of the scientific consensus, and aked if they accept it. Of particular interest were people who are low in TiS, and CTists. Our main hypotheses were:

**H1: Higher trust in science is associated with more science knowledge and more acceptance of the scientific consensus**

**H2: Higher conspiracy thinking/belief is associated with less science knowledge and less acceptance of the scientific consensus**

In study 1, we plainly inform participants of the scientific consensus and ask if they accept it. We use a subset of questions on basic science knowledge that have been used in numerous public opinion surveys [@allumScienceKnowledgeAttitudes2008; @durantPublicUnderstandingScience1989; @millerMeasurementCivicScientific1998a] sometimes referred to as the "Oxford scale" [@gauchatCulturalAuthorityScience2011]. In study 2, we remove a problematic science question and present people with a more elaborated explanation of the scientific consensus, as well as additional sources. This allows us to remove the issue that participants might simply not have trusted us to correctly report the scientific consensus in study 1. Study 3 follows essentially the same design as study 2, but is run on a vaccine-skeptic sample. Further, we provided participants with an explicit option to revise their answer, in case they rejected the consensus. We also asked follow-up questions on why participants accepted the scientific consensus, namely whether they thought it was because they trust scientists or because they verified independently. Study 4 uses the same design as study 3, but using a set of more recent, advance science questions that participants are less likely to have encountered before, and that they would be less likely to be able to verify (or even understand) themselves. Just as study 3, study 4 is run on a sample of vaccine-skeptic participants.

# Methods

```{r exp1}
# Analyze data of experiments and store results

# Experiment 1

# read the cleaned long version data set
exp1_long <- read_csv("exp_1/data/cleaned_long.csv")

# read wide version data set
exp1_wide <- read_csv("exp_1/data/cleaned_wide.csv")

# Research questions

# RQ3
cor_trust_knowledge <- cor.test(exp1_wide$wgm_sciencegeneral, exp1_wide$avg_knowledge) %>% text_ready()

cor_trust_acceptance <- cor.test(exp1_wide$wgm_sciencegeneral, exp1_wide$avg_acceptance) %>% text_ready()

# RQ4
cor_conspiracy_knowledge <- cor.test(exp1_wide$BCTI_avg, exp1_wide$avg_knowledge) %>% text_ready()

cor_conspiracy_acceptance <- cor.test(exp1_wide$BCTI_avg, exp1_wide$avg_acceptance)%>% text_ready()

# Exploratory: Acceptance and trust in science/conspiracy thinking conditional on false answers only
exp1_false_knowledge <- exp1_long %>% 
  # make numeric versions
  mutate(acceptance_num = ifelse(acceptance == "Yes", 1, 0)
  ) %>% 
  group_by(id, knowledge) %>% 
  # calculate by-participant averages
  summarize(
    n = n(),
    n_accepted = sum(acceptance_num),
    avg_acceptance  = sum(acceptance_num)/n(), 
    wgm_sciencegeneral = mean(wgm_sciencegeneral), 
    BCTI_avg = mean(BCTI_avg), 
    # add additonal measures for robustness checks
    CMQ_avg = mean(CMQ_avg), 
    SICBS = mean(SICBS),
    wgm_scientists = mean(wgm_scientists), 
    pew = mean(pew)
  ) %>% 
  # filter to only false responses
  filter(knowledge == FALSE) %>% 
  ungroup()

false_answers_cor_trust_acceptance <- cor.test(exp1_false_knowledge$wgm_sciencegeneral, exp1_false_knowledge$avg_acceptance) %>% text_ready()

false_answers_cor_conspiracy_acceptance <- cor.test(exp1_false_knowledge$BCTI_avg, exp1_false_knowledge$avg_acceptance) %>% text_ready()

# extract descriptives for inline reporting
exp1_descriptives <- list(
  # Demographics
  n_subj = n_distinct(exp1_wide$id),
  gender = exp1_wide %>% group_by(gender) %>% summarize(n = n_distinct(id)) %>% split(.$gender),
  age = exp1_wide %>% summarize(across(age,list(mean = mean, median = median, sd = sd), 
                                       .names = "{.fn}")) %>% rounded_numbers(),
  # RQ1 & RQ2
  means = exp1_wide %>% 
    summarize(across(c(avg_knowledge, avg_acceptance), 
                     list(mean = mean, sd = sd), 
                     .names = "{.col}_{.fn}")) %>%
    rename_with(~sub("^avg_", "", .), everything()) %>% 
    mutate_if(is.numeric, round, digits = 2) %>% 
  mutate(across(ends_with("_mean"), ~paste0(.*100, " %"))) ,
  # RQ3
  cor_trust_knowledge = cor_trust_knowledge,
  cor_trust_acceptance = cor_trust_acceptance,
  # RQ4
  cor_conspiracy_knowledge = cor_conspiracy_knowledge,
  cor_conspiracy_acceptance = cor_conspiracy_acceptance, 
  # Exploratory
  # acceptance and knowledge
  conditional_acceptance = exp1_long %>% 
    group_by(knowledge, acceptance) %>% 
    count() %>% 
    group_by(knowledge) %>% 
    mutate (share = n/sum(n), 
            # rename knowledge values
            knowledge = ifelse(knowledge == FALSE, "false", "true")) %>% 
    mutate_if(is.numeric, round, digits = 3) %>% 
    mutate (share = paste0(share*100, " %")) %>% 
    super_split(knowledge, acceptance),
  # acceptance and trust in science/conspiracy thinking
  false_answers_cor_trust_acceptance = false_answers_cor_trust_acceptance,
  false_answers_cor_conspiracy_acceptance = false_answers_cor_conspiracy_acceptance
)
```

```{r exp2}
# Analyze data of experiments and store results

# Experiment 2

# read the cleaned long version data set
exp2_long <- read_csv("exp_2/data/cleaned_long.csv")

# read wide version data set
exp2_wide <- read_csv("exp_2/data/cleaned_wide.csv")

# read coded justifications
exp2_justifications <- read_csv("exp_2/data/justifications_clean.csv")

# Research questions

# H1a
cor_trust_knowledge <- cor.test(exp2_wide$wgm_sciencegeneral, exp2_wide$avg_knowledge) %>% text_ready()

# H2a
cor_conspiracy_knowledge <- cor.test(exp2_wide$BCTI_avg, exp2_wide$avg_knowledge) %>% text_ready()

# Exploratory: Acceptance and trust in science/conspiracy thinking conditional on false answers only
exp2_false_knowledge <- exp2_long %>% 
  # make numeric versions
  mutate(acceptance_num = ifelse(acceptance == "Yes", 1, 0)
  ) %>% 
  group_by(id, knowledge) %>% 
  # calculate by-participant averages
  summarize(
    n = n(),
    n_accepted = sum(acceptance_num),
    avg_acceptance  = sum(acceptance_num)/n(), 
    wgm_sciencegeneral = mean(wgm_sciencegeneral), 
    BCTI_avg = mean(BCTI_avg), 
    # add additonal measures for robustness checks
    CMQ_avg = mean(CMQ_avg), 
    SICBS = mean(SICBS),
    wgm_scientists = mean(wgm_scientists), 
    pew = mean(pew)
  ) %>% 
  # filter to only false responses
  filter(knowledge == FALSE) %>% 
  ungroup()

# H1b
# conditional on false answers
false_answers_cor_trust_acceptance <- cor.test(exp2_false_knowledge$wgm_sciencegeneral, exp2_false_knowledge$avg_acceptance) %>% text_ready()
# pooled across all cases (in line with all other studies' analysis)
cor_trust_acceptance <- cor.test(exp2_wide$wgm_sciencegeneral, exp2_wide$avg_acceptance) %>% text_ready()

# H2b
# conditional on false answers
false_answers_cor_conspiracy_acceptance <- cor.test(exp2_false_knowledge$BCTI_avg, exp2_false_knowledge$avg_acceptance) %>% text_ready()
# pooled across all cases (in line with all other studies' analysis)
cor_conspiracy_acceptance <- cor.test(exp2_wide$BCTI_avg, exp2_wide$avg_acceptance) %>% text_ready()

# extract descriptives for inline reporting
exp2_descriptives <- list(
  # Demographics
  n_subj = n_distinct(exp2_wide$id),
  gender = exp2_wide %>% group_by(gender) %>% summarize(n = n_distinct(id)) %>% split(.$gender),
  age = exp2_wide %>% summarize(across(age,list(mean = mean, median = median, sd = sd), 
                                       .names = "{.fn}")) %>% rounded_numbers(),
  # RQ1 & RQ2
  means = exp2_wide %>% 
    summarize(across(c(avg_knowledge, avg_acceptance), 
                     list(mean = mean, sd = sd), 
                     .names = "{.col}_{.fn}")) %>%
    rename_with(~sub("^avg_", "", .), everything()) %>% 
    mutate_if(is.numeric, round, digits = 2) %>% 
  mutate(across(ends_with("_mean"), ~paste0(.*100, " %"))) ,
  # H1
  # a
  cor_trust_knowledge = cor_trust_knowledge,
  # b (conditional on false responses)
  false_answers_cor_trust_acceptance = false_answers_cor_trust_acceptance,
  # b (all cases)
  cor_trust_acceptance = cor_trust_acceptance,
  # H2
  # a
  cor_conspiracy_knowledge = cor_conspiracy_knowledge,
  # b (conditional on false responses)
  false_answers_cor_conspiracy_acceptance = false_answers_cor_conspiracy_acceptance,
  # b(all cases)
  cor_conspiracy_acceptance = cor_conspiracy_acceptance, 
  # Exploratory
  # acceptance and knowledge
  conditional_acceptance = exp2_long %>% 
    group_by(knowledge, acceptance) %>% 
    count() %>% 
    group_by(knowledge) %>% 
    mutate (share = n/sum(n), 
            # rename knowledge values
            knowledge = ifelse(knowledge == FALSE, "false", "true")) %>% 
    mutate_if(is.numeric, round, digits = 3) %>% 
    mutate (share = paste0(share*100, " %")) %>% 
    super_split(knowledge, acceptance),
  # Justifications
  justifications_n = nrow(exp2_justifications), 
  justifications_n_participants = n_distinct(exp2_justifications$id),
  justification_by_category = exp2_justifications %>%
    group_by(category)%>%
    summarize(n = n(), 
              n_subjects = n_distinct(id)) %>%
    mutate(share = n/sum(n)) %>%
    mutate_if(is.numeric, round, digits=3) %>% 
    mutate(share = paste0(share*100, "%")) %>% 
    split(.$category)
)
```

```{r exp3}
# Analyze data of experiments and store results

# Experiment 3

# read the cleaned long version data set
exp3_long <- read_csv("exp_3/data/cleaned_long.csv")

# read wide version data set
exp3_wide <- read_csv("exp_3/data/cleaned_wide.csv")

# read coded justifications
exp3_justifications <- read_csv("exp_3/data/justifications_clean.csv")

# Research questions

# H1
cor_trust_knowledge <- cor.test(exp3_wide$wgm_sciencegeneral, exp3_wide$avg_knowledge,
  use = "complete.obs") %>% text_ready()
cor_trust_acceptance <- cor.test(exp3_wide$wgm_sciencegeneral, exp3_wide$avg_acceptance,
  use = "complete.obs") %>% text_ready()

# H2
cor_conspiracy_knowledge <- cor.test(exp3_wide$BCTI_avg, exp3_wide$avg_knowledge,
  use = "complete.obs") %>% text_ready()
cor_conspiracy_acceptance <- cor.test(exp3_wide$BCTI_avg, exp3_wide$avg_acceptance,
  use = "complete.obs") %>% text_ready()

# Exploratory: Acceptance and trust in science/conspiracy thinking conditional on false answers only
exp3_false_knowledge <- exp3_long %>% 
  # make numeric versions
  mutate(acceptance_num = ifelse(acceptance == "Yes", 1, 0)
  ) %>% 
  group_by(id, knowledge) %>% 
  # calculate by-participant averages
  summarize(
    n = n(),
    n_accepted = sum(acceptance_num),
    avg_acceptance  = sum(acceptance_num)/n(), 
    wgm_sciencegeneral = mean(wgm_sciencegeneral), 
    BCTI_avg = mean(BCTI_avg), 
    # add additonal measures for robustness checks
    CMQ_avg = mean(CMQ_avg), 
    SICBS = mean(SICBS),
    wgm_scientists = mean(wgm_scientists), 
    pew = mean(pew)
  ) %>% 
  # filter to only false responses
  filter(knowledge == FALSE) %>% 
  ungroup()

# H1b conditional on false answers
false_answers_cor_trust_acceptance <- cor.test(exp3_false_knowledge$wgm_sciencegeneral, exp3_false_knowledge$avg_acceptance,
  use = "complete.obs") %>% 
  text_ready()

# H2b conditional on false answers
false_answers_cor_conspiracy_acceptance <- cor.test(exp3_false_knowledge$BCTI_avg, exp3_false_knowledge$avg_acceptance,
  use = "complete.obs") %>% 
  text_ready()

# extract descriptives for inline reporting
exp3_descriptives <- list(
  # Demographics
  n_subj = n_distinct(exp3_wide$id),
  n_subject_by_outcome = exp3_wide %>% 
    summarize(across(c("wgm_sciencegeneral", "reason_agreement", "BCTI_avg"), 
                     ~sum(!is.na(.x)
                     )
    )
    ),
  gender = exp3_wide %>% group_by(gender) %>% summarize(n = n_distinct(id)) %>% split(.$gender),
  age = exp3_wide %>% summarize(across(age,list(mean = mean, median = median, sd = sd), 
                                       .names = "{.fn}")) %>% rounded_numbers(),
  # RQ1 & RQ2
  means = exp3_wide %>% 
    summarize(across(c(avg_knowledge, avg_acceptance, avg_acceptance_initial), 
                     list(mean = mean, sd = sd), 
                     .names = "{.col}_{.fn}")) %>%
    rename_with(~sub("^avg_", "", .), everything()) %>% 
    mutate_if(is.numeric, round, digits = 2) %>% 
  mutate(across(ends_with("_mean"), ~paste0(.*100, " %"))) ,
  # H1
  # a
  cor_trust_knowledge = cor_trust_knowledge,
  # b (conditional on false responses)
  false_answers_cor_trust_acceptance = false_answers_cor_trust_acceptance,
  # b (all cases)
  cor_trust_acceptance = cor_trust_acceptance,
  # H2
  # a
  cor_conspiracy_knowledge = cor_conspiracy_knowledge,
  # b (conditional on false responses)
  false_answers_cor_conspiracy_acceptance = false_answers_cor_conspiracy_acceptance,
  # b(all cases)
  cor_conspiracy_acceptance = cor_conspiracy_acceptance, 
  # Exploratory
  # acceptance and knowledge
  conditional_acceptance = exp3_long %>% 
    pivot_longer(c(acceptance, acceptance_initial), 
                 names_to = "measure", 
                 values_to = "acceptance") %>% 
    group_by(knowledge, measure, acceptance) %>% 
    count() %>% 
    group_by(knowledge, measure) %>% 
    mutate (share = n/sum(n), 
            # rename knowledge values
            knowledge = ifelse(knowledge == FALSE, "false", "true")) %>% 
    mutate_if(is.numeric, round, digits = 3) %>% 
    mutate (share = paste0(share*100, " %")) %>% 
    super_split(knowledge,  measure, acceptance),
  # Justifications
  justifications_n = nrow(exp3_justifications), 
  justifications_n_participants = n_distinct(exp3_justifications$id),
  justification_by_category = exp3_justifications %>%
    group_by(category)%>%
    summarize(n = n(), 
              n_subjects = n_distinct(id)) %>%
    mutate(share = n/sum(n)) %>%
    mutate_if(is.numeric, round, digits=3) %>% 
    mutate(share = paste0(share*100, "%")) %>% 
    split(.$category),
  # Consensus acceptance
  acceptance_n = exp3_wide %>% 
    filter(!is.na(reason_agreement)) %>% 
    nrow(.),
  acceptance_by_reason = exp3_wide %>% 
    group_by(reason_agreement) %>% 
    summarize(n = n()) %>%
    drop_na(reason_agreement) %>% 
    mutate(share = n/sum(n)) %>%
    mutate_if(is.numeric, round, digits=3) %>% 
    mutate(share = paste0(share*100, "%")) %>% 
    split(.$reason_agreement),
  reason_followup_n = exp3_wide %>% 
    filter(!is.na(reason_followup)) %>% 
    nrow(.)
)

```

```{r exp4}
# Analyze data of experiments and store results

# Experiment 4

# read the cleaned long version data set
exp4_long <- read_csv("exp_4/data/cleaned_long.csv")

# read wide version data set
exp4_wide <- read_csv("exp_4/data/cleaned_wide.csv")

# read coded justifications
exp4_justifications <- read_csv("exp_4/data/justifications_clean.csv")

# Research questions

# H1
cor_trust_knowledge <- cor.test(exp4_wide$wgm_sciencegeneral, exp4_wide$avg_knowledge,
  use = "complete.obs") %>% text_ready()
cor_trust_acceptance <- cor.test(exp4_wide$wgm_sciencegeneral, exp4_wide$avg_acceptance,
  use = "complete.obs") %>% text_ready()

# H2
cor_conspiracy_knowledge <- cor.test(exp4_wide$BCTI_avg, exp4_wide$avg_knowledge,
  use = "complete.obs") %>% text_ready()
cor_conspiracy_acceptance <- cor.test(exp4_wide$BCTI_avg, exp4_wide$avg_acceptance,
  use = "complete.obs") %>% text_ready()

# Exploratory: Acceptance and trust in science/conspiracy thinking conditional on false answers only
exp4_false_knowledge <- exp4_long %>% 
  # make numeric versions
  mutate(acceptance_num = ifelse(acceptance == "Yes", 1, 0)
  ) %>% 
  group_by(id, knowledge) %>% 
  # calculate by-participant averages
  summarize(
    n = n(),
    n_accepted = sum(acceptance_num),
    avg_acceptance  = sum(acceptance_num)/n(), 
    wgm_sciencegeneral = mean(wgm_sciencegeneral), 
    BCTI_avg = mean(BCTI_avg), 
    # add additonal measures for robustness checks
    CMQ_avg = mean(CMQ_avg), 
    SICBS = mean(SICBS),
    wgm_scientists = mean(wgm_scientists), 
    pew = mean(pew)
  ) %>% 
  # filter to only false responses
  filter(knowledge == FALSE) %>% 
  ungroup()

# H1b conditional on false answers
false_answers_cor_trust_acceptance <- cor.test(exp4_false_knowledge$wgm_sciencegeneral, exp4_false_knowledge$avg_acceptance,
  use = "complete.obs") %>% 
  text_ready()

# H2b conditional on false answers
false_answers_cor_conspiracy_acceptance <- cor.test(exp4_false_knowledge$BCTI_avg, exp4_false_knowledge$avg_acceptance,
  use = "complete.obs") %>% 
  text_ready()

# extract descriptives for inline reporting
exp4_descriptives <- list(
  # Demographics
  n_subj = n_distinct(exp4_wide$id),
  n_subject_by_outcome = exp4_wide %>% 
    summarize(across(c("wgm_sciencegeneral", "reason_agreement", "BCTI_avg"), 
                     ~sum(!is.na(.x)
                     )
    )
    ),
  gender = exp4_wide %>% group_by(gender) %>% summarize(n = n_distinct(id)) %>% split(.$gender),
  age = exp4_wide %>% summarize(across(age,list(mean = mean, median = median, sd = sd), 
                                       .names = "{.fn}")) %>% rounded_numbers(),
  # RQ1 & RQ2
  means = exp4_wide %>% 
    summarize(across(c(avg_knowledge, avg_acceptance, avg_acceptance_initial), 
                     list(mean = mean, sd = sd), 
                     .names = "{.col}_{.fn}")) %>%
    rename_with(~sub("^avg_", "", .), everything()) %>% 
    mutate_if(is.numeric, round, digits = 2) %>% 
  mutate(across(ends_with("_mean"), ~paste0(.*100, " %"))) ,
  # H1
  # a
  cor_trust_knowledge = cor_trust_knowledge,
  # b (conditional on false responses)
  false_answers_cor_trust_acceptance = false_answers_cor_trust_acceptance,
  # b (all cases)
  cor_trust_acceptance = cor_trust_acceptance,
  # H2
  # a
  cor_conspiracy_knowledge = cor_conspiracy_knowledge,
  # b (conditional on false responses)
  false_answers_cor_conspiracy_acceptance = false_answers_cor_conspiracy_acceptance,
  # b(all cases)
  cor_conspiracy_acceptance = cor_conspiracy_acceptance, 
  # Exploratory
  # acceptance and knowledge
  conditional_acceptance = exp4_long %>% 
    pivot_longer(c(acceptance, acceptance_initial), 
                 names_to = "measure", 
                 values_to = "acceptance") %>% 
    group_by(knowledge, measure, acceptance) %>% 
    count() %>% 
    group_by(knowledge, measure) %>% 
    mutate (share = n/sum(n), 
            # rename knowledge values
            knowledge = ifelse(knowledge == FALSE, "false", "true")) %>% 
    mutate_if(is.numeric, round, digits = 3) %>% 
    mutate (share = paste0(share*100, " %")) %>% 
    super_split(knowledge,  measure, acceptance),
  # Justifications
  justifications_n = nrow(exp4_justifications), 
  justifications_n_participants = n_distinct(exp4_justifications$id),
  justification_by_category = exp4_justifications %>%
    group_by(category)%>%
    summarize(n = n(), 
              n_subjects = n_distinct(id)) %>%
    mutate(share = n/sum(n)) %>%
    mutate_if(is.numeric, round, digits=3) %>% 
    mutate(share = paste0(share*100, "%")) %>% 
    split(.$category),
  # Consensus acceptance
  acceptance_n = exp4_wide %>% 
    filter(!is.na(reason_agreement)) %>% 
    nrow(.),
  acceptance_by_reason = exp4_wide %>% 
    group_by(reason_agreement) %>% 
    summarize(n = n()) %>%
    drop_na(reason_agreement) %>% 
    mutate(share = n/sum(n)) %>%
    mutate_if(is.numeric, round, digits=3) %>% 
    mutate(share = paste0(share*100, "%")) %>% 
    split(.$reason_agreement),
  reason_followup_n = exp4_wide %>% 
    filter(!is.na(reason_followup)) %>% 
    nrow(.)
)

```

## Procedure

After providing their consent to participate in the study, participants were given an attention check "While watching the television, have you ever had a fatal heart attack?" [1-6; 1 = Never, 6 = Often]. All participants who did not answer "1 = Never" were excluded. Participants then read the following instructions:"We will ask you 10 questions about science. After each question, we will provide you with the scientifically consensual answer and ask whether you accept it." Next, participants answered a set of 10 basic science questions (in study 1 they were randomly selected from a pool of 11 questions) in random order. After each question, participants were presented with an answer reflecting the scientific consensus. In study 2 and 3, participants additionally saw a short explanation which we wrote, partly based on explanations generated by ChatGPT, and three links to authoritative sources supporting the answer. In study 4, we provided only two links and no explanation. Participants were asked to choose whether they accept the answer or not, before proceeding to the next question. Participants then answered questions on conspiracy thinking, conspiracy beliefs and trust in science.

In study 2, 3 and 4, we presented participants with open-ended questions to explain cases where they rejected the scientific consensus. In study 3 and 4, we additionally gave participants the option to change their answer and accept the scientific consensus. Finally, at the end of study 3 and 4, we asked participants: "For the questions in which you agreed with the scientific consensus, would you say that...?" The answer options were: (i) "You mostly agree with the consensus because, on that question, you trust scientists", (ii) "You mostly agree with the consensus because you have been able to independently verify it", and (iii) "Other", with a text box for participants to explain. Participants who selected “You mostly agree with the consensus because you have been able to independently verify it”, were asked the open-ended follow-up question: "Could you please tell us how you independently verified the information?".

## Participants

We recruited a total of `r exp1_descriptives$n_subj + exp2_descriptives$n_subj + exp3_descriptives$n_subj + exp4_descriptives$n_subj` (`r exp1_descriptives$n_subj` in study 1, `r exp2_descriptives$n_subj` in study 2, `r exp3_descriptives$n_subj` in study 3, `r exp4_descriptives$n_subj` in study 4) participants in the US via prolific. Details and demographics can be found in the online supplemental material. While samples for study 1 and 2 were convenience samples, study 3 and 4 were run on a sample holding vaccine-skeptic beliefs (see OSM for details).

## Materials

Table \@ref(tab:knowledge) shows all questions and their answer options used in the four studies.

```{r knowledge}
# Function to replace special characters in all cells of a data frame
replace_special_characters <- function(df) {
  df %>%
    mutate(across(everything(), ~ gsub("\u2013|\u2014|\u2212", "-", .))) %>% # Replace en-dash, em-dash, and Unicode minus with ASCII hyphen-minus
    mutate(across(everything(), ~ iconv(., from = "UTF-8", to = "ASCII//TRANSLIT"))) # Ensure all text is ASCII
}

# Read the CSV file
items <- read_csv("exp_4/materials/overview_questions.csv")

# Replace special characters
items_cleaned <- replace_special_characters(items) %>% 
  mutate(id = 1:nrow(.)) %>%
  select(id, everything())  # Ensure 'id' is the first column

# add footnote for the tree items
items_cleaned[11, 2] <- paste0("*", 
                               items_cleaned[11, 2])

# Output the table
knitr::kable(items_cleaned, booktabs = T, longtable = TRUE,
    caption = "Science knowledge items", 
    full_width = T, 
    col.names = c("", "Study 1-3", "Study 4")) %>%
  kable_styling(font_size = 8) %>%
  column_spec(1, width = "2em") %>%
  column_spec(2, width = "22em")%>%
  column_spec(3, width = "22em") %>% 
  footnote(symbol = "Only used in Study 1")
```

Our main measure of conspiracy belief was asking participants about specific conspiracy theories. We selected 10 science/health related conspiracy theories from the Belief in Conspiracy Theory Inventory (BCTI) [@pennycookOverconfidentlyConspiratorialConspiracy2022] (Table \@ref(tab:conspiracy)). Participants were asked: "Below is a list of events for which the official version has been disputed. For each event, we would like you to indicate to what extent you believe the cover-up version of events is true or false. [1-9; labels: 1 - completly false, 5 - unsure, 9 - completely true]".

As a robustness check, we also used established scales on conspiracy thinking, namely the four-item conspiracy mentality questionnaire (CMQ) [@bruderMeasuringIndividualDifferences2013] and the single item conspiracy beliefs scale (SICBS) [@lantianMeasuringBeliefConspiracy2016]. Details and comparisons between the scales can be found in the OSM.

```{r conspiracy, echo=FALSE}
# Create the data frame
items <- c(
  "The Apollo moon landings never happened and were staged in a Hollywood film studio.",
  "A cure for cancer was discovered years ago, but this has been suppressed by the pharmaceutical industry and the U.S. Food and Drug Administration (FDA).",
  "The spread of certain viruses and/or diseases is the result of the deliberate, concealed efforts of vested interests.",
  "The claim that the climate is changing due to emissions from fossil fuels is a hoax perpetrated by corrupt scientists who want to spend more taxpayer money on climate research.",
  "The Earth is flat (not spherical) and this fact has been covered up by scientists and vested interests.",
  "There is a causal link between vaccination and autism that has been covered up by the pharmaceutical industry.",
  "In the 1950s and 1960s more than 100 million Americans received a polio vaccine contaminated with a potentially cancer-causing virus.",
  "Proof of alien contact is being concealed from the public.",
  "Hydroxychloroquine has been demonstrated to be a safe and effective treatment of COVID and this information is being suppressed.",
  "Dinosaurs never existed, evolution is not real, and scientists have been faking the fossil record.")

conspiracy_items <- data.frame(id = 1:length(items), items = items)

# Output the table
kbl(conspiracy_items, booktabs = T, longtable = T, col.names = NULL, 
    caption = "Conspiracy items", 
    full_width = F) %>%
  kable_styling(font_size = 8) %>% 
  column_spec(1) %>%
  column_spec(2, width = "40em")
```

Our main item for measuring trust in science was selected from the Wellcome Global Monitor survey: "In general, would you say that you trust science a lot, some, not much, or not at all? [1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]". We also included two additional trust questions, one also from the Wellcome Global Monitor (WGM) survey ("How much do you trust scientists in this country? Do you trust them a lot, some, not much, or not at all? [1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]"), the other from the Pew research center ("How much confidence do you have in scientists to act in the best interests of the public? [1-5; 1 = No confidence at all, 5 = A great deal of confidence]"). We selected these items so that we could compare the ratings in our sample to global survey results. The WGM survey has been administered in over 140 countries and included over 140000 respondents. The Pew question has recently been used by a world-wide many labs study in 67 countries with 71417 respondents [@colognaTrustScientistsTheir2024].

# Results

(ref:summary-plot) Points represent the average share of acceptance as a function of the **A** level of trust in science ("In general, would you say that you trust science a lot, some, not much, or not at all? [1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]"); **B** average conspiracy thinking (CMQ, 5 items on a scale from 0 to 100); **C** belief in a specific conspiracy theory. Participants were asked to rate their belief in the conspiracy on a scale from 1 to 9, with the labels: "1 - completly false, 5 - unsure, 9 - completely true". Averages are only based on participants who believed at least to some degree in the conspiracy theory. We consider a believer as anyone scoring higher than the scale midpoint, i.e. \>5. Number labels in plots represent the number of participants for the corresponding data point.

```{r summary-plot, fig.cap="(ref:summary-plot)", fig.height= 10, fig.width=10}
# combine data of all three studies
combined_data <- bind_rows(exp1_long %>% 
                             mutate(study = "study 1"), 
                           exp2_long %>% 
                             mutate(study = "study 2"), 
                           exp3_long %>% 
                             mutate(study = "study 3"),
                           exp4_long %>% 
                             mutate(study = "study 4")
                           )

# make plot data for trust
plot_trust <- combined_data %>% 
  drop_na(wgm_sciencegeneral, acceptance) %>% 
  # make a labels version of trust in science
  mutate(wgm_sciencegeneral = case_when(
    wgm_sciencegeneral == 1 ~ "1 (Not at all)",
    wgm_sciencegeneral == 2 ~ "2 (Not much)",
    wgm_sciencegeneral == 3 ~ "3 (Some)",
    wgm_sciencegeneral == 4 ~ "4 (A lot)",
    TRUE ~ as.character(wgm_sciencegeneral)  # Handle any other cases (optional)
  )) %>% 
  group_by(study, wgm_sciencegeneral, acceptance) %>% 
  summarize(n = n(), 
            n_participants = n_distinct(id)) %>% 
  group_by(study, wgm_sciencegeneral) %>% 
  mutate (share = n/sum(n), 
          n_participants = max(n_participants)) %>% 
  mutate_if(is.numeric, round, digits = 3)

# make trust plot
trust <- ggplot(plot_trust %>% 
         filter(acceptance == "Yes"), 
       aes(x = wgm_sciencegeneral, y = share, color = study, shape = study)) +
  geom_point(position = "dodge", alpha = 0.5) +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1), limits = c(0, 1)) +
  geom_text_repel(aes(label = paste0(n_participants)),
            vjust = -0.5, size = 3,
            show.legend = FALSE) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  labs(x = "Trust in science", y = "Acceptance rate of consensus", color = "Study", shape = "Study") +
  plot_theme + 
  theme(legend.position = "top")

# make plot data for conspiracy belief
plot_conspiracy <- combined_data %>% 
  # bring to long format
  pivot_longer(cols = BCTI_apollo:BCTI_evolution, 
               names_to = "BCTI_item", values_to = "score") %>% 
  # make a belief variable if participant indicates a numeric answer above the
  # scale midpoint
  mutate(belief = ifelse(score > 5, TRUE, 
                         ifelse(is.na(score), NA, FALSE)
                         ),
         BCTI_item = sub("^BCTI_", "", BCTI_item)
         ) %>% 
  group_by(study, BCTI_item, belief, acceptance) %>% 
  summarize(n = n(),
            n_participants = n_distinct(id)) %>% 
  group_by(study, BCTI_item, belief) %>% 
  mutate (share = n/sum(n), 
          n_participants = max(n_participants)) %>% 
  mutate_if(is.numeric, round, digits = 3) %>% 
  ungroup()

# make conspiracy belief plot
conspiracy <- ggplot(plot_conspiracy %>% 
         filter(acceptance == "Yes" & belief == TRUE), 
       aes(x = BCTI_item, y = share, color = study, shape = study)) +
  geom_point(position = "dodge", alpha = 0.5) +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1), limits = c(0, 1)) +
  geom_text_repel(aes(label = paste0(n_participants)),
            vjust = -0.5, size = 3,
            show.legend = FALSE) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  labs(x = "Conspiracy theory belief", y = "Acceptance rate of consensus", color = "Study", shape = "Study") +
  plot_theme +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

# Conspiracy Thinking

# Define breaks for each integer from 1 to 9
breaks <- seq(0, 100, by = 20)

# Define labels for each bin in the desired format
labels <- paste0(c(0, 20, 40, 60, 80), "-", c(20, 40, 60, 80, 100))

plot_data <- combined_data %>% 
  drop_na(CMQ_avg) %>% 
  # Create the binned version of CMQ_avg
  mutate(CMQ_avg_binned = cut(CMQ_avg, breaks = breaks, labels = labels, include.lowest = TRUE, right = FALSE)) %>% 
  group_by(study, CMQ_avg_binned, acceptance) %>% 
  summarize(n = n()) %>% 
  group_by(study, CMQ_avg_binned) %>% 
  mutate (share = n/sum(n)) %>% 
  mutate_if(is.numeric, round, digits = 3)

n_participants <- combined_data %>% 
  drop_na(CMQ_avg) %>% 
  # Create the binned version of CMQ_avg
  mutate(CMQ_avg_binned = cut(CMQ_avg, breaks = breaks, labels = labels, include.lowest = TRUE, right = FALSE)) %>% 
  group_by(study, CMQ_avg_binned) %>% 
  summarize(n_participants = n_distinct(id))

plot_conspiracy_thinking <- left_join(plot_data, n_participants)

# make plot
conspiracy_thinking  <- ggplot(plot_conspiracy_thinking  %>% 
         filter(acceptance == "Yes"), 
       aes(x = CMQ_avg_binned, y = share, color = study, shape = study)) +
  geom_point(position = "dodge", alpha = 0.5) +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1), limits = c(0, 1)) +
  geom_text_repel(aes(label = paste0(n_participants)),
            vjust = -0.5, size = 3,
            show.legend = FALSE) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  labs(x = "Conspiracy Thinking", y = "Acceptance rate of consensus", color = "Study", shape = "Study") +
  plot_theme + 
  theme(legend.position = "top")

# remove elements of plots for better overall plot
conspiracy_thinking <- conspiracy_thinking + 
  #rremove("ylab") +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) 
  #guides(color = "none", shape = "none")

conspiracy <- conspiracy +
  rremove("ylab") +
  guides(color = "none", shape = "none")

trust <- trust +
  rremove("ylab")

# common plot
patchwork <- (trust + conspiracy_thinking)/conspiracy + 
  plot_layout(guides = 'collect', widths = c(1, 1)) +
  plot_annotation(tag_levels = 'A')  & theme(legend.position = "top")

wrap_elements(patchwork) +
  labs(tag = "Acceptance rate of consensus") +
    theme(
    plot.tag = element_text(angle = 90, , face = "bold", size = 12),
    plot.tag.position = "left"
  )
```

Across all studies we find that almost all participants trust basic and recent/advance science facts almost all the time. That is true even for participants who expressed low general trust in science, and those who have a tendency to conspiracy think or hold science-related conspiracy beliefs (Fig. \@ref(fig:summary-plot)).

Our results suggest that the more people tend to trust science, the more they tend to know about science and accept the scientific consensus. In study 2 and 3, we find a positive but small correlation between trust in science and both science knowledge (study 1: r = `r exp1_descriptives$cor_trust_knowledge$estimate`, p `r exp1_descriptives$cor_trust_knowledge$p.value`; study 2: r = `r exp2_descriptives$cor_trust_knowledge$estimate`, p `r exp2_descriptives$cor_trust_knowledge$p.value`), and acceptance of scientific consensus (study 1: r = `r exp1_descriptives$cor_trust_acceptance$estimate`, p `r exp1_descriptives$cor_trust_acceptance$p.value`; study 2: r = `r exp2_descriptives$cor_trust_acceptance$estimate`, p `r exp2_descriptives$cor_trust_acceptance$p.value`). In study 3 and 4--both run on vaccine-skeptic samples--we find no statistically significant correlation between science knowledge and trust in science (study 3: r = `r exp3_descriptives$cor_trust_knowledge$estimate`, p `r exp3_descriptives$cor_trust_knowledge$p.value`; study 4: r = `r exp4_descriptives$cor_trust_knowledge$estimate`, p `r exp4_descriptives$cor_trust_knowledge$p.value`), but a small positive correlation between acceptance of scientific consensus and trust in science (study 3: r = `r exp3_descriptives$cor_trust_acceptance$estimate`, p `r exp3_descriptives$cor_trust_acceptance$p.value`; study 4: r = `r exp4_descriptives$cor_trust_acceptance$estimate`, p `r exp4_descriptives$cor_trust_acceptance$p.value`).

As for conspiracy beliefs, results are mixed. In studies 1 and 2, we find a small negative correlation between conspiracy thinking and both science knowledge (study 1: r = `r exp1_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp1_descriptives$cor_conspiracy_knowledge$p.value`; study 2: r = `r exp2_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp2_descriptives$cor_conspiracy_knowledge$p.value`), and acceptance of scientific consensus (study 1: r = `r exp1_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp1_descriptives$cor_conspiracy_acceptance$p.value`; study 2: r = `r exp2_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp2_descriptives$cor_conspiracy_acceptance$p.value`). However, in study 3 and 4, we do not find a statistically significant correlation with neither science knowledge (study 3: r = `r exp3_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp3_descriptives$cor_conspiracy_knowledge$p.value`; study 4: r = `r exp4_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp4_descriptives$cor_conspiracy_knowledge$p.value`), and between nor acceptance of scientific consensus (study 3: r = `r exp3_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp3_descriptives$cor_conspiracy_acceptance$p.value`; study 4: r = `r exp4_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp4_descriptives$cor_conspiracy_acceptance$p.value`).

Overall, the correlations we find are relatively weak, which might be partly due to ceiling effects. This was the case in particular for acceptance: In all studies, in the vast majority of cases, participants accepted accepted science facts. The absence of correlations between conspiracy beliefs and science knowledge or acceptance in studies 3 and 4 might be explained by our sample selection. For these studies, we recruited samples scoring considerably higher in conspiracy belief than the convenience samples of study 1 and 2 (see OSM). The presence of effects in study 1 and 2, and the absence in study 3 and 4 might be because differences between moderate and high conspiracy thinking might not matter, while, differences between low and moderate/high conspiracy thinking do matter.

Did participants who had initially gotten it wrong change their minds towards the scientific consensus? Yes. In most cases (study 1: `r exp1_descriptives$conditional_acceptance$false$Yes$share`; study 2: `r exp2_descriptives$conditional_acceptance$false$Yes$share`; study 3: `r exp3_descriptives$conditional_acceptance$false$acceptance$Yes$share`; study 4: `r exp4_descriptives$conditional_acceptance$false$acceptance$Yes$share`), participants readily accepted the scientific consensus after having initially given the wrong answer to a question. Only in study 4 do we find evidence[^3] that changing one's mind towards the scientific consensus is associated with (more) trust in science (study 1: r = `r exp1_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp1_descriptives$false_answers_cor_trust_acceptance$p.value`; study 2: r = `r exp2_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp2_descriptives$false_answers_cor_trust_acceptance$p.value`; study 3: r = `r exp3_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp3_descriptives$false_answers_cor_trust_acceptance$p.value`; study 4: r = `r exp4_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp4_descriptives$false_answers_cor_trust_acceptance$p.value`) and only in study 2 evidence that it is associated with (less) conspiracy thinking (study 1: r = `r exp1_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp1_descriptives$false_answers_cor_conspiracy_acceptance$p.value`; study 2: r = `r exp2_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp2_descriptives$false_answers_cor_conspiracy_acceptance$p.value`; study 3: r = `r exp3_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp3_descriptives$false_answers_cor_conspiracy_acceptance$p.value`; study 4: r = `r exp4_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp4_descriptives$false_answers_cor_conspiracy_acceptance$p.value`). In very few cases, participants changed their mind away from the scientific consensus. These participants initially gave the correct response, but rejected the scientific consensus right after, thereby contradicting their own initial response (study 1: `r exp1_descriptives$conditional_acceptance$true$No$share`; study 2: `r exp2_descriptives$conditional_acceptance$true$No$share`; study 3: `r exp3_descriptives$conditional_acceptance$true$acceptance$No$share`; study 4: `r exp4_descriptives$conditional_acceptance$true$acceptance$No$share`).

[^3]: For this analysis, we restricted our data to those cases in which participants initially gave the wrong answer.

Why did participants reject the scientific consensus? We got a total of `r exp4_descriptives$justifications_n + exp3_descriptives$justifications_n + exp2_descriptives$justifications_n` answers (study 2: `r exp2_descriptives$justifications_n`; study 3: `r exp3_descriptives$justifications_n`; study 4: `r exp4_descriptives$justifications_n`) from `r exp4_descriptives$justifications_n_participants + exp3_descriptives$justifications_n_participants + exp2_descriptives$justifications_n_participants` (study 2: `r exp2_descriptives$justifications_n_participants`; study 3: `r exp3_descriptives$justifications_n_participants`; study 4: `r exp4_descriptives$justifications_n_participants`) different participants to the open-ended questions on why they had rejected the scientific consensus on a particular question. Based on the answers, we created five categories. Table \@ref(tab:justifications) summarizes these answers by five categories. All answers can be read in the OSM.

```{r justifications}

## make a common table of exp 2 and exp 3
justifications <- bind_rows(exp2_justifications %>% 
                              mutate(study = "study 2"), 
                            exp3_justifications %>% 
                              mutate(study = "study 3", 
                                     final_coding = as.character(final_coding)),
                            exp4_justifications %>% 
                              mutate(study = "study 4"))

justifications %>%
  group_by(category)%>%
  summarize(n = n(), 
            n_subjects = n_distinct(id)) %>%
  mutate(Share = n/sum(n)) %>%
  mutate_if(is.numeric, round, digits=3) %>% 
  mutate(Share = paste0(Share*100, "%")) %>% 
  arrange(desc(n)) %>% 
  select(category, n, Share, n_subjects) %>% 
  ungroup() %>% 
  rename(
    Category = category,
    `N (instances)` = n, 
    `Share (instances)` = Share, 
    `N (unique participants)` = n_subjects
  ) %>% 
  apa_table(caption = "Justifications by category for data from studies 2, 3 and 4 combined.")
```

Why did participants accept the scientific consensus? In studies 3 and 4, we had asked participants about cases where they agreed with the scientific consensus. A total of `r exp3_descriptives$acceptance_n + exp4_descriptives$acceptance_n` (study 3: `r exp3_descriptives$acceptance_n`; study 4: `r exp4_descriptives$acceptance_n`) participants answered this question. Of these, `r exp3_descriptives$acceptance_by_reason[["trust in scientists"]]$share` in study 3 and `r exp4_descriptives$acceptance_by_reason[["trust in scientists"]]$share` in study 4 said they accepted the scientific consensus because they trust scientists, while `r exp3_descriptives$acceptance_by_reason[["independent verification"]]$share` in study 3 and `r exp4_descriptives$acceptance_by_reason[["independent verification"]]$share` in study 4 said they independently verified the fact. We asked all participants who gave this answer to explain how they did it (answers can be found in the OSM). `r exp3_descriptives$acceptance_by_reason[["other"]]$share` in study 3 and `r exp4_descriptives$acceptance_by_reason[["other"]]$share` in study 4 answered with other "other" and gave an open-ended explanation (see OSM).

More detailed results addressing all our pre-registered research questions can be found in the OSM.

# Discussion

Survey data from the US suggests a decline of trust in science in recent years [@brianAmericansTrustScientists2023]. Across the globe, many people only trust science ‘some of the time’ , and a sizable minority doesn’t trust it [@wellcomeglobalmonitorWellcomeGlobalMonitor2018; @wellcomeglobalmonitorWellcomeGlobalMonitor2020]. This appears difficult to reconcile with the fact that, in many countries, most people nowadays receive at least a basic education [@ritchieGlobalEducation2023], and that science education has been found to be the most strongly associated variable with trust in science [@wellcomeglobalmonitorWellcomeGlobalMonitor2018].

But how deep is distrust of science? In four studies, we have shown that almost every participant accepts almost all of basic science questions we included. Even when people do not know the correct answer to a science question, they tend to mostly accept the scientific consensus afterwards. That is true not only for basic science questions (studies 1-3), but also for more recent scientific findings that most people are unfamiliar with (study 4). In study 1 the acceptance rate after having given a wrong answer was considerably lower (`r exp1_descriptives$conditional_acceptance$false$No$share`) than in subsequent studies (study 2: `r exp2_descriptives$conditional_acceptance$false$Yes$share`; study 3: `r exp3_descriptives$conditional_acceptance$false$acceptance$Yes$share`; study 4: `r exp4_descriptives$conditional_acceptance$false$acceptance$Yes$share`). While this could be just sampling variation, it might be that adding explanations and/or sources convinced participants more than merely stating the consensus. In general, participants with lower trust in science and who believe more in conspiracy theories tended to both know less about science and accept the scientific consensus less.

Scholars have long interpreted the lack of basic science knowledge as the root cause of distrust in and more generally unfavorable attitudes towards science, an explanation known under the term "deficit-model" [@millerMeasurementCivicScientific1998a; @millerPublicUnderstandingAttitudes2004; @durantPublicUnderstandingScience1989]. In line with proponents of the deficit-model, we find that people indeed err even on very basic science questions. However, we also find that almost everyone instantly accepted the scientific consensus once informed about it. Moreover, the same was true for non-basic science questions. This calls into question the deficit model: If participants trust science on basic (and non-basic) questions—as suggests immediate acceptance of the scientific consensus—then unfavorable attitudes towards science must have other explanations than a mere lack of knowledge. This argument is supported by past research. In a seminal meta-analysis, \@allumScienceKnowledgeAttitudes2008 found correlations between science knowledge and general attitudes towards science to be at best small and largely context dependent [\@allumScienceKnowledgeAttitudes2008]. A more recent report by the National Academies of Sciences, Engineering, and Medicine states that science knowledge is "unlikely to substantially affect attitudes on scientific issues" [@nationalacademiesofsciencesCommunicatingScienceEffectively2017].

How to make sense of the detachment of basic science knowledge from attitudes towards science? One explanation might be that most scientific findings are very abstract. For example, to most people, it will not be relevant for their daily lives to know that the speed of light is 299'792'458 meters per second. It is therefor easy to simply accept it, all the more if it appears to be consensual within the population of scientists. In other words, people might hold on to most scientific findings as reflective beliefs—beliefs that have little effect on our behaviors [@mercierHowGoodAre2021]. It is probably when it becomes tangible that science shape's people's attitudes. This can happen, for example, when science appears to support conclusions in favor of certain policies or moral positions, as for example in the case of anthropogenic climate change or vaccines.

It is tempting, especially for scientists, to explain mistrust in science with ignorance. After all, careful proof of objective truth is what makes science trustworthy—that is, for scientists. For most people, what shapes trust in science might not be so different from what shapes trust in other institutions, such as the government, banks or church.

\FloatBarrier

# References

::: {#refs}
:::

\newpage

# (APPENDIX) Appendix {.unnumbered}

```{r child = "appendix_exp1.Rmd"}
```

\clearpage

```{r child = "appendix_exp2.Rmd"}
```

\clearpage

```{r child = "appendix_exp3.Rmd"}
```

\clearpage

```{r child = "appendix_exp4.Rmd"}
```