### 2024-09-19 appendix_knowledge.Rmd

# Science knowledge {#knowledge}

Figure \@ref(fig:knowledge-overview) shows the distributions of basic science knowledge across the four studies.

(ref:knowledge-overview) Distributions of science knowledge.

```{r knowledge-overview, fig.cap="(ref:knowledge-overview)", fig.width=8, fig.height=7}
plot_data <- combined_data %>%
  mutate(study = paste0("Study ", study)) %>% 
  group_by(study, id) %>% 
  summarize(n_correct = sum(knowledge)) %>% 
  group_by(study, n_correct) %>% 
  summarize(n = n_distinct(id)) %>% 
  group_by(study) %>% 
  mutate(share = n/sum(n)) %>% 
  rounded_numbers() %>% 
  ungroup()

ggplot(plot_data, aes(x = n_correct, y = share)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1)) +
  scale_x_continuous(breaks = seq(0, 10, 1)) +
  labs(x = "Number of correct answers (out of 10 possible)", y = "Share of participants") +
  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
            position = position_dodge(width = 0.9),
            vjust = -0.5, size = 3) +
  plot_theme +
  facet_wrap(~study)
```