### 2024-09-19 paper.Rmd

---
title             : "Quasi-universal acceptance of basic science in the US"
shorttitle        : "Does anyone mistrust basic science?"

header-includes:  | # to prevent floats from moving past certain points (for the appendix)
  \usepackage{placeins}

author: 
  - name          : ""
    affiliation   : ""

affiliation:
  - id            : ""
    institution   : ""

abstract: |
  Substantial minorities of the population report a low degree of trust in science, or endorse conspiracy theories that violate basic scientific knowledge. This might indicate a wholesale rejection of science. In four studies, we asked 782 US participants questions about trust in science, conspiracy beliefs, and basic science (e.g. the relative size of electrons and atoms). Participants were provided with the scientifically consensual answer to the basic science questions, and asked whether they accept it. Acceptance of the scientific consensus was very high in the sample as a whole (95.1%), but also in every sub-sample (e.g., no trust in science: 87.3%; complete endorsement of flat Earth theory: 87.2%). This quasi-universal acceptance of basic science suggests that people are motivated to reject specific scientific beliefs, and not science as a whole. This could be leveraged in science communication.

keywords          : Trust in science; conspiracy theories; conspiracy thinking; science knowledge.

floatsintext      : yes
linenumbers       : no 
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "doc" # "doc" for nice look, "man" for manuscripty
output            : papaja::apa6_pdf # "_doc" for word; however, note that some of the features of kableExtra are not available for word and will yield errors. For example to knit to word, you'll have to comment out all "add_header_above()" functions for model output tables, or e.g. set always allow html as true in the yaml heading

always_allow_html: true

appendix:
  - "appendix_exp1.Rmd"
  - "appendix_exp2.Rmd"
  - "appendix_exp3.Rmd"
  - "appendix_exp4.Rmd"
  - "appendix_acceptance_knowledge.Rmd"
  - "appendix_conspiracy_thinking.Rmd"
  - "appendix_trust.Rmd"
  - "appendix_conspiracy_belief.Rmd"
  - "appendix_knowledge.Rmd"
  - "appendix_by-question.Rmd"

bibliography: references.bib
---

```{r setup, include=FALSE}
# Figure out output format
is_docx <- knitr::pandoc_to("docx") | knitr::pandoc_to("odt")
is_latex <- knitr::pandoc_to("latex")
is_html <- knitr::pandoc_to("html")

# Word-specific things
table_format <- ifelse(is_docx, "huxtable", "kableExtra")  # Huxtable tables
conditional_dpi <- ifelse(is_docx, 300, 300)  # Higher DPI
conditional_align <- ifelse(is_docx, "default", "center")  # Word doesn't support align

# Knitr options
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  # tidy.opts = list(width.cutoff = 120),  # Code width
  # fig.retina = 3, dpi = conditional_dpi,
  # fig.width = 7, fig.asp = 0.618,
  # fig.align = conditional_align, out.width = "100%",
  fig.path = "output/figures/",
  cache.path = "output/_cache/",
  fig.process = function(x) {  # Remove "-1" from figure names
    x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
    if (file.rename(x, x2)) x2 else x
  },
  options(scipen = 99999999)  # Prevent scientific notation
)

# R options
options(
  width = 90,  # Output width
  dplyr.summarise.inform = FALSE,  # Turn off dplyr's summarize() auto messages
  knitr.kable.NA = "",  # Make NAs blank in kables
  kableExtra.latex.load_packages = FALSE,  # Don't add LaTeX preamble stuff
  modelsummary_factory_default = table_format,  # Set modelsummary backend
  modelsummary_format_numeric_latex = "plain"  # Don't use siunitx
)
```

```{r packages, include=FALSE}
# load required packages
library("papaja")      # For APA style manuscript   
library("lme4")        # model specification / estimation
library("lmerTest")    # provides p-values in the output
library("tidyverse")   # data wrangling and visualisation
library("afex")        # anova and deriving p-values from lmer
library("broom")       # extracting data from model fits 
library("broom.mixed") # extracting data from mixed models
library("metafor")     # doing mata analysis
library("patchwork")   # put several plots together
library("ggridges")    # for plots
library("gghalves")    # for plots
library("ggbeeswarm")  # Special distribution-shaped point jittering
library("knitr")       # for tables
library("kableExtra")  # also for tables
library("ggpubr")      # for combining plots with ggarrange() 
library("grid")        # for image plots   
library("gridExtra")   # for image plots
library("png")         # for image plots
library("modelsummary") # for regression tables
library("ggrepel") # for better random patterns in plots
```

```{r functions}
# load plot theme
source("functions/plot_theme.R")

# load other functions
source("functions/own_functions.R")
```

# Introduction

Trust in science is related to many desirable outcomes, from acceptance of anthropogenic climate change [@colognaRoleTrustClimate2020] or vaccination [@sturgisTrustScienceSocial2021; @lindholtPublicAcceptanceCOVID192021] to following recommendations during COVID [@alganTrustScientistsTimes2021, which suggests that trust in science was the most important predictor of these behaviors].

Unfortunately, people who report a high degree of trust in science are a minority in most countries, and they are outnumbered by people who have low trust in science in many areas, e.g., most of Africa and significant parts of Asia [@wellcomeglobalmonitorWellcomeGlobalMonitor2018; @wellcomeglobalmonitorWellcomeGlobalMonitor2020]. Moreover, trust in science has recently been declining in some countries [@alganTrustScientistsTimes2021; although see @wellcomeglobalmonitorPublicTrustScientists2021], including the US [@brianAmericansTrustScientists2023], where it is also increasingly polarizing [@gauchatPoliticizationSciencePublic2012; @krauseTrendsAmericansTrust2019a; @liPolarizationPublicTrust2022]. Rejection of science can take the more extreme form of conspiracy theories, many of which question the scientific consensus on vaccination, climate change, and even the shape of the Earth.

What does this apparent rejection of science actually entail? Do people who say they do not trust science, or who believe in conspiracy theories at odds with well-established science, reject most of science? Or, on the contrary, do they object to a few specific facets of science, while still accepting the overwhelming majority of basic science? Answering this question has theoretical implications: if some people reject science wholesale, then it’s possible that their distrust of science causes their rejection of specific scientific theories, or facilitates their acceptance of conspiracy theories; by contrast, if everyone trusts most of science, then the mistrust of specific facets of science is more likely to be post-hoc, a rationalization of other beliefs or behaviors. The present question also has practical implications: many communication attempts leverage the scientific consensus [e.g. on vaccination, climate change, etc., for review, see @vanstekelenburgScientificConsensusCommunicationContested2022; see also @veckalov27countryTestCommunicating2024]. These attempts are more likely to be successful if everyone trusts basic science than if some people reject science wholesale.

Studies of science knowledge provide relevant evidence, as people who answer science knowledge questions correctly presumably trust scientists in this respect. The average level of science knowledge is quite low, even in countries with extensive science education. For example, only 55% in the US (in 2014) and only 46% in the EU (in 2005) knew that antibiotics kill only bacteria and not viruses [@nationalacademiesofsciencesengineeringandmedicineScienceLiteracyConcepts2016]. However, these results only provide a lower bound on how much people trust basic science, as participants who do not provide the correct answer might accept it when it is pointed out to them.

## The present studies

In a series of four pre-registered online studies (total n = 782), we asked US participants questions about well-established scientific facts. For each question, we asked participants what they thought the correct answer was (testing their knowledge of science), we informed them of the scientifically accepted answer, and asked them whether they accepted it (measuring their trust in basic science). We also measured participants’ trust in science using standard measures, as well as their beliefs in various conspiracy theories and their tendency to engage in conspiratorial thinking. The four studies, including materials, hypotheses, and analyses, were pre-registered and all materials and data are accessible via the Open Science Framework (OSF). The differences between the four studies are summarized presently, and the methods are detailed below.

### Materials

In Study 1, we used questions drawn from questionnaires of scientific knowledge (e.g. "Are electrons smaller, larger, or the same size as atoms? [Smaller; Same size; Larger]"), supplemented by a ‘trick’ question ("Where do trees mainly draw the materials with which they create their mass? [Earth; Water; Air]"; correct answer: Air). In Studies 2 and 3, this last question was removed. The scientific facts used in Studies 1 to 3 represent long-established and basic knowledge. In Study 4 we used more recent much less basic scientific discoveries (e.g. "What is the electric charge of the Higgs Boson, as established in 2012? [1.602176634 × 10-19; 0; 3.2×10-19C]"; correct answer: 0, i.e. electrically neutral).

### Presentation of the scientific consensus

In Study 1, we simply told participants that they would be provided with the scientifically consensual answer. However, for participants to accept this answer, they must not only trust science, but also trust that we are presenting them with the actual scientifically consensual answer. To remove this issue, in Studies 2 to 3, we presented participants with a short explanation of the correct answer, as well as links to three sources per answer (e.g. Wikipedia, National Geographic or NASA). In Study 4, as the topics were more complex, we did not provide an explanation, but still provided two sources per answer.

### Measure of acceptance of the scientific consensus

In Study 1, we simply look at whether participants say that they accept the scientifically consensual answer. In the subsequent studies, we asked participants to explain why they disagreed with the scientific consensus. This revealed that a number of participants had made a mistake (misunderstanding, selecting the wrong answer). As a result, in Studies 3 and 4, participants who have indicated that they rejected the scientifically consensual answer were offered the option to revise their answer, or to keep rejecting it.

### Additional questions

In Studies 3 and 4, we attempted to understand how some people who say they do not trust science still accept scientifically consensual answers, by asking them whether they accepted the answers on the basis of trust in science or because they had independently verified them.

### Samples

Studies 1 and 2 were conducted on the standard sample of US participants recruited on the platform Prolific Academic. In order to increase the share of participants with low trust in science, and who endorse conspiracy theories, Studies 3 and 4 used the same platform, but only recruited participants who had declared previously being skeptical of vaccination.

### Hypotheses

The main goal of the present studies is descriptive: to find out whether participants who report not trusting science, or who believe in conspiracy theories, still accept most well-established scientific facts. However, we also tested two directional hypotheses (pre-registered as research questions in the Study 1):

**H1: Higher trust in science is associated with more science knowledge and more acceptance of the scientific consensus**

**H2: Higher conspiracy thinking/belief is associated with less science knowledge and less acceptance of the scientific consensus**

# Methods

```{r exp1}
# Analyze data of experiments and store results

# Experiment 1

# read the cleaned long version data set
exp1_long <- read_csv("exp_1/data/cleaned_long.csv")

# read wide version data set
exp1_wide <- read_csv("exp_1/data/cleaned_wide.csv")

# Research questions

# RQ3
cor_trust_knowledge <- cor.test(exp1_wide$wgm_sciencegeneral, exp1_wide$avg_knowledge) %>% text_ready()

cor_trust_acceptance <- cor.test(exp1_wide$wgm_sciencegeneral, exp1_wide$avg_acceptance) %>% text_ready()

# RQ4
cor_conspiracy_knowledge <- cor.test(exp1_wide$BCTI_avg, exp1_wide$avg_knowledge) %>% text_ready()

cor_conspiracy_acceptance <- cor.test(exp1_wide$BCTI_avg, exp1_wide$avg_acceptance)%>% text_ready()

# Exploratory: Acceptance and trust in science/conspiracy thinking conditional on false answers only
exp1_false_knowledge <- exp1_long %>% 
  # make numeric versions
  mutate(acceptance_num = ifelse(acceptance == "Yes", 1, 0)
  ) %>% 
  group_by(id, knowledge) %>% 
  # calculate by-participant averages
  summarize(
    n = n(),
    n_accepted = sum(acceptance_num),
    avg_acceptance  = sum(acceptance_num)/n(), 
    wgm_sciencegeneral = mean(wgm_sciencegeneral), 
    BCTI_avg = mean(BCTI_avg), 
    # add additonal measures for robustness checks
    CMQ_avg = mean(CMQ_avg), 
    SICBS = mean(SICBS),
    wgm_scientists = mean(wgm_scientists), 
    pew = mean(pew)
  ) %>% 
  # filter to only false responses
  filter(knowledge == FALSE) %>% 
  ungroup()

false_answers_cor_trust_acceptance <- cor.test(exp1_false_knowledge$wgm_sciencegeneral, exp1_false_knowledge$avg_acceptance) %>% text_ready()

false_answers_cor_conspiracy_acceptance <- cor.test(exp1_false_knowledge$BCTI_avg, exp1_false_knowledge$avg_acceptance) %>% text_ready()

# extract descriptives for inline reporting
exp1_descriptives <- list(
  # Demographics
  n_subj = n_distinct(exp1_wide$id),
  gender = exp1_wide %>% group_by(gender) %>% summarize(n = n_distinct(id)) %>% split(.$gender),
  age = exp1_wide %>% summarize(across(age,list(mean = mean, median = median, sd = sd), 
                                       .names = "{.fn}")) %>% rounded_numbers(),
  # RQ1 & RQ2
  means = exp1_wide %>% 
    summarize(across(c(avg_knowledge, avg_acceptance), 
                     list(mean = mean, sd = sd), 
                     .names = "{.col}_{.fn}")) %>%
    rename_with(~sub("^avg_", "", .), everything()) %>% 
    mutate_if(is.numeric, round, digits = 2) %>% 
    mutate(across(ends_with("_mean"), ~paste0(.*100, " %"))),
  means_numeric = exp1_wide %>% 
    summarize(across(c(avg_knowledge, avg_acceptance,), 
                     list(mean = mean, sd = sd), 
                     .names = "{.col}_{.fn}")) %>%
    rename_with(~sub("^avg_", "", .), everything()),
  # RQ3
  cor_trust_knowledge = cor_trust_knowledge,
  cor_trust_acceptance = cor_trust_acceptance,
  # RQ4
  cor_conspiracy_knowledge = cor_conspiracy_knowledge,
  cor_conspiracy_acceptance = cor_conspiracy_acceptance, 
  # Exploratory
  # acceptance and knowledge
  conditional_acceptance = exp1_long %>% 
    group_by(knowledge, acceptance) %>% 
    count() %>% 
    group_by(knowledge) %>% 
    mutate (share = n/sum(n), 
            # rename knowledge values
            knowledge = ifelse(knowledge == FALSE, "false", "true")) %>% 
    mutate_if(is.numeric, round, digits = 3) %>% 
    mutate (share = paste0(share*100, " %")) %>% 
    super_split(knowledge, acceptance),
  # acceptance and trust in science/conspiracy thinking
  false_answers_cor_trust_acceptance = false_answers_cor_trust_acceptance,
  false_answers_cor_conspiracy_acceptance = false_answers_cor_conspiracy_acceptance
)
```

```{r exp2}
# Analyze data of experiments and store results

# Experiment 2

# read the cleaned long version data set
exp2_long <- read_csv("exp_2/data/cleaned_long.csv")

# read wide version data set
exp2_wide <- read_csv("exp_2/data/cleaned_wide.csv")

# read coded justifications
exp2_justifications <- read_csv("exp_2/data/justifications_clean.csv")

# Research questions

# H1a
cor_trust_knowledge <- cor.test(exp2_wide$wgm_sciencegeneral, exp2_wide$avg_knowledge) %>% text_ready()

# H2a
cor_conspiracy_knowledge <- cor.test(exp2_wide$BCTI_avg, exp2_wide$avg_knowledge) %>% text_ready()

# Exploratory: Acceptance and trust in science/conspiracy thinking conditional on false answers only
exp2_false_knowledge <- exp2_long %>% 
  # make numeric versions
  mutate(acceptance_num = ifelse(acceptance == "Yes", 1, 0)
  ) %>% 
  group_by(id, knowledge) %>% 
  # calculate by-participant averages
  summarize(
    n = n(),
    n_accepted = sum(acceptance_num),
    avg_acceptance  = sum(acceptance_num)/n(), 
    wgm_sciencegeneral = mean(wgm_sciencegeneral), 
    BCTI_avg = mean(BCTI_avg), 
    # add additonal measures for robustness checks
    CMQ_avg = mean(CMQ_avg), 
    SICBS = mean(SICBS),
    wgm_scientists = mean(wgm_scientists), 
    pew = mean(pew)
  ) %>% 
  # filter to only false responses
  filter(knowledge == FALSE) %>% 
  ungroup()

# H1b
# conditional on false answers
false_answers_cor_trust_acceptance <- cor.test(exp2_false_knowledge$wgm_sciencegeneral, exp2_false_knowledge$avg_acceptance) %>% text_ready()
# pooled across all cases (in line with all other studies' analysis)
cor_trust_acceptance <- cor.test(exp2_wide$wgm_sciencegeneral, exp2_wide$avg_acceptance) %>% text_ready()

# H2b
# conditional on false answers
false_answers_cor_conspiracy_acceptance <- cor.test(exp2_false_knowledge$BCTI_avg, exp2_false_knowledge$avg_acceptance) %>% text_ready()
# pooled across all cases (in line with all other studies' analysis)
cor_conspiracy_acceptance <- cor.test(exp2_wide$BCTI_avg, exp2_wide$avg_acceptance) %>% text_ready()

# extract descriptives for inline reporting
exp2_descriptives <- list(
  # Demographics
  n_subj = n_distinct(exp2_wide$id),
  gender = exp2_wide %>% group_by(gender) %>% summarize(n = n_distinct(id)) %>% split(.$gender),
  age = exp2_wide %>% summarize(across(age,list(mean = mean, median = median, sd = sd), 
                                       .names = "{.fn}")) %>% rounded_numbers(),
  # RQ1 & RQ2
  means = exp2_wide %>% 
    summarize(across(c(avg_knowledge, avg_acceptance), 
                     list(mean = mean, sd = sd), 
                     .names = "{.col}_{.fn}")) %>%
    rename_with(~sub("^avg_", "", .), everything()) %>% 
    mutate_if(is.numeric, round, digits = 2) %>% 
    mutate(across(ends_with("_mean"), ~paste0(.*100, " %"))) ,
  means_numeric = exp2_wide %>% 
    summarize(across(c(avg_knowledge, avg_acceptance,), 
                     list(mean = mean, sd = sd), 
                     .names = "{.col}_{.fn}")) %>%
    rename_with(~sub("^avg_", "", .), everything()),
  # H1
  # a
  cor_trust_knowledge = cor_trust_knowledge,
  # b (conditional on false responses)
  false_answers_cor_trust_acceptance = false_answers_cor_trust_acceptance,
  # b (all cases)
  cor_trust_acceptance = cor_trust_acceptance,
  # H2
  # a
  cor_conspiracy_knowledge = cor_conspiracy_knowledge,
  # b (conditional on false responses)
  false_answers_cor_conspiracy_acceptance = false_answers_cor_conspiracy_acceptance,
  # b(all cases)
  cor_conspiracy_acceptance = cor_conspiracy_acceptance, 
  # Exploratory
  # acceptance and knowledge
  conditional_acceptance = exp2_long %>% 
    group_by(knowledge, acceptance) %>% 
    count() %>% 
    group_by(knowledge) %>% 
    mutate (share = n/sum(n), 
            # rename knowledge values
            knowledge = ifelse(knowledge == FALSE, "false", "true")) %>% 
    mutate_if(is.numeric, round, digits = 3) %>% 
    mutate (share = paste0(share*100, " %")) %>% 
    super_split(knowledge, acceptance),
  # Justifications
  justifications_n = nrow(exp2_justifications), 
  justifications_n_participants = n_distinct(exp2_justifications$id),
  justification_by_category = exp2_justifications %>%
    group_by(category)%>%
    summarize(n = n(), 
              n_subjects = n_distinct(id)) %>%
    mutate(share = n/sum(n)) %>%
    mutate_if(is.numeric, round, digits=3) %>% 
    mutate(share = paste0(share*100, "%")) %>% 
    split(.$category)
)
```

```{r exp3}
# Analyze data of experiments and store results

# Experiment 3

# read the cleaned long version data set
exp3_long <- read_csv("exp_3/data/cleaned_long.csv")

# read wide version data set
exp3_wide <- read_csv("exp_3/data/cleaned_wide.csv")

# read coded justifications
exp3_justifications <- read_csv("exp_3/data/justifications_clean.csv")

# Research questions

# H1
cor_trust_knowledge <- cor.test(exp3_wide$wgm_sciencegeneral, exp3_wide$avg_knowledge,
  use = "complete.obs") %>% text_ready()
cor_trust_acceptance <- cor.test(exp3_wide$wgm_sciencegeneral, exp3_wide$avg_acceptance,
  use = "complete.obs") %>% text_ready()

# H2
cor_conspiracy_knowledge <- cor.test(exp3_wide$BCTI_avg, exp3_wide$avg_knowledge,
  use = "complete.obs") %>% text_ready()
cor_conspiracy_acceptance <- cor.test(exp3_wide$BCTI_avg, exp3_wide$avg_acceptance,
  use = "complete.obs") %>% text_ready()

# Exploratory: Acceptance and trust in science/conspiracy thinking conditional on false answers only
exp3_false_knowledge <- exp3_long %>% 
  # make numeric versions
  mutate(acceptance_num = ifelse(acceptance == "Yes", 1, 0)
  ) %>% 
  group_by(id, knowledge) %>% 
  # calculate by-participant averages
  summarize(
    n = n(),
    n_accepted = sum(acceptance_num),
    avg_acceptance  = sum(acceptance_num)/n(), 
    wgm_sciencegeneral = mean(wgm_sciencegeneral), 
    BCTI_avg = mean(BCTI_avg), 
    # add additonal measures for robustness checks
    CMQ_avg = mean(CMQ_avg), 
    SICBS = mean(SICBS),
    wgm_scientists = mean(wgm_scientists), 
    pew = mean(pew)
  ) %>% 
  # filter to only false responses
  filter(knowledge == FALSE) %>% 
  ungroup()

# H1b conditional on false answers
false_answers_cor_trust_acceptance <- cor.test(exp3_false_knowledge$wgm_sciencegeneral, exp3_false_knowledge$avg_acceptance,
  use = "complete.obs") %>% 
  text_ready()

# H2b conditional on false answers
false_answers_cor_conspiracy_acceptance <- cor.test(exp3_false_knowledge$BCTI_avg, exp3_false_knowledge$avg_acceptance,
  use = "complete.obs") %>% 
  text_ready()

# Exploratory: Differences by reason

# trust
trust_by_reason_model <- run_regression(exp3_wide, dependent_var = "wgm_sciencegeneral", independent_var = "reason_agreement")

# acceptance
acceptance_by_reason_model <-run_regression(exp3_wide, dependent_var = "avg_acceptance", independent_var = "reason_agreement")

# BCTI (conspiracy beliefs)
BCTI_by_reason_model <- run_regression(exp3_wide, dependent_var = "BCTI_avg", independent_var = "reason_agreement")

# survey time
time_by_reason_model <- run_regression(exp3_wide %>% filter(duration_mins < 30), dependent_var = "duration_mins", independent_var = "reason_agreement")

# extract descriptives for inline reporting
exp3_descriptives <- list(
  # Demographics
  n_subj = n_distinct(exp3_wide$id),
  n_subject_by_outcome = exp3_wide %>% 
    summarize(across(c("wgm_sciencegeneral", "reason_agreement", "BCTI_avg"), 
                     ~sum(!is.na(.x)
                     )
    )
    ),
  gender = exp3_wide %>% group_by(gender) %>% summarize(n = n_distinct(id)) %>% split(.$gender),
  age = exp3_wide %>% summarize(across(age,list(mean = mean, median = median, sd = sd), 
                                       .names = "{.fn}")) %>% rounded_numbers(),
  # RQ1 & RQ2
  means = exp3_wide %>% 
    summarize(across(c(avg_knowledge, avg_acceptance, avg_acceptance_initial), 
                     list(mean = mean, sd = sd), 
                     .names = "{.col}_{.fn}")) %>%
    rename_with(~sub("^avg_", "", .), everything()) %>% 
    mutate_if(is.numeric, round, digits = 2) %>% 
  mutate(across(ends_with("_mean"), ~paste0(.*100, " %"))) ,
  means_numeric = exp3_wide %>% 
    summarize(across(c(avg_knowledge, avg_acceptance, avg_acceptance_initial), 
                     list(mean = mean, sd = sd), 
                     .names = "{.col}_{.fn}")) %>%
    rename_with(~sub("^avg_", "", .), everything()),
  # H1
  # a
  cor_trust_knowledge = cor_trust_knowledge,
  # b (conditional on false responses)
  false_answers_cor_trust_acceptance = false_answers_cor_trust_acceptance,
  # b (all cases)
  cor_trust_acceptance = cor_trust_acceptance,
  # H2
  # a
  cor_conspiracy_knowledge = cor_conspiracy_knowledge,
  # b (conditional on false responses)
  false_answers_cor_conspiracy_acceptance = false_answers_cor_conspiracy_acceptance,
  # b(all cases)
  cor_conspiracy_acceptance = cor_conspiracy_acceptance, 
  # Exploratory
  # acceptance and knowledge
  conditional_acceptance = exp3_long %>% 
    pivot_longer(c(acceptance, acceptance_initial), 
                 names_to = "measure", 
                 values_to = "acceptance") %>% 
    group_by(knowledge, measure, acceptance) %>% 
    count() %>% 
    group_by(knowledge, measure) %>% 
    mutate (share = n/sum(n), 
            # rename knowledge values
            knowledge = ifelse(knowledge == FALSE, "false", "true")) %>% 
    mutate_if(is.numeric, round, digits = 3) %>% 
    mutate (share = paste0(share*100, " %")) %>% 
    super_split(knowledge,  measure, acceptance),
  # Justifications
  justifications_n = nrow(exp3_justifications), 
  justifications_n_participants = n_distinct(exp3_justifications$id),
  justification_by_category = exp3_justifications %>%
    group_by(category)%>%
    summarize(n = n(), 
              n_subjects = n_distinct(id)) %>%
    mutate(share = n/sum(n)) %>%
    mutate_if(is.numeric, round, digits=3) %>% 
    mutate(share = paste0(share*100, "%")) %>% 
    split(.$category),
  # Consensus acceptance
  acceptance_n = exp3_wide %>% 
    filter(!is.na(reason_agreement)) %>% 
    nrow(.),
  acceptance_by_reason = exp3_wide %>% 
    group_by(reason_agreement) %>% 
    summarize(n = n()) %>%
    drop_na(reason_agreement) %>% 
    mutate(share = n/sum(n)) %>%
    mutate_if(is.numeric, round, digits=3) %>% 
    mutate(share = paste0(share*100, "%")) %>% 
    split(.$reason_agreement),
  reason_followup_n = exp3_wide %>% 
    filter(!is.na(reason_followup)) %>% 
    nrow(.),
  # Exploratory: Differences by reason for trust
  reason_means = exp3_wide %>% 
    group_by(reason_agreement) %>% 
    summarize(across(c(wgm_sciencegeneral, avg_acceptance, BCTI_avg), 
                     list(mean = ~mean(.x, na.rm = TRUE), 
                          sd = ~sd(.x, na.rm = TRUE)), 
                     .names = "{.col}_{.fn}")) %>%
    rename_with(~sub("^avg_", "", .), everything()) %>% 
    mutate_if(is.numeric, round, digits = 2) %>% 
    mutate(across(acceptance_mean, ~paste0(.*100, " %"))) %>% 
    split(.$reason_agreement), 
  reason_trust_model = trust_by_reason_model %>% split(.$term), 
  reason_acceptance_model = acceptance_by_reason_model %>% split(.$term),
  reason_BCTI_model = BCTI_by_reason_model %>%  split(.$term),
  reason_time_model = time_by_reason_model %>%  split(.$term), 
  n_participants_over30mins = exp3_wide %>% 
    filter(duration_mins >= 30) %>% 
    nrow(.)
)
```

```{r exp4}
# Analyze data of experiments and store results

# Experiment 4

# read the cleaned long version data set
exp4_long <- read_csv("exp_4/data/cleaned_long.csv")

# read wide version data set
exp4_wide <- read_csv("exp_4/data/cleaned_wide.csv")

# read coded justifications
exp4_justifications <- read_csv("exp_4/data/justifications_clean.csv")

# Research questions

# H1
cor_trust_knowledge <- cor.test(exp4_wide$wgm_sciencegeneral, exp4_wide$avg_knowledge,
  use = "complete.obs") %>% text_ready()
cor_trust_acceptance <- cor.test(exp4_wide$wgm_sciencegeneral, exp4_wide$avg_acceptance,
  use = "complete.obs") %>% text_ready()

# H2
cor_conspiracy_knowledge <- cor.test(exp4_wide$BCTI_avg, exp4_wide$avg_knowledge,
  use = "complete.obs") %>% text_ready()
cor_conspiracy_acceptance <- cor.test(exp4_wide$BCTI_avg, exp4_wide$avg_acceptance,
  use = "complete.obs") %>% text_ready()

# Exploratory: Acceptance and trust in science/conspiracy thinking conditional on false answers only
exp4_false_knowledge <- exp4_long %>% 
  # make numeric versions
  mutate(acceptance_num = ifelse(acceptance == "Yes", 1, 0)
  ) %>% 
  group_by(id, knowledge) %>% 
  # calculate by-participant averages
  summarize(
    n = n(),
    n_accepted = sum(acceptance_num),
    avg_acceptance  = sum(acceptance_num)/n(), 
    wgm_sciencegeneral = mean(wgm_sciencegeneral), 
    BCTI_avg = mean(BCTI_avg), 
    # add additonal measures for robustness checks
    CMQ_avg = mean(CMQ_avg), 
    SICBS = mean(SICBS),
    wgm_scientists = mean(wgm_scientists), 
    pew = mean(pew)
  ) %>% 
  # filter to only false responses
  filter(knowledge == FALSE) %>% 
  ungroup()

# H1b conditional on false answers
false_answers_cor_trust_acceptance <- cor.test(exp4_false_knowledge$wgm_sciencegeneral, exp4_false_knowledge$avg_acceptance,
  use = "complete.obs") %>% 
  text_ready()

# H2b conditional on false answers
false_answers_cor_conspiracy_acceptance <- cor.test(exp4_false_knowledge$BCTI_avg, exp4_false_knowledge$avg_acceptance,
  use = "complete.obs") %>% 
  text_ready()

# Exploratory: Differences by reason

# trust
trust_by_reason_model <- run_regression(exp4_wide, dependent_var = "wgm_sciencegeneral", independent_var = "reason_agreement")

# acceptance
acceptance_by_reason_model <-run_regression(exp4_wide, dependent_var = "avg_acceptance", independent_var = "reason_agreement")

# BCTI (conspiracy beliefs)
BCTI_by_reason_model <- run_regression(exp4_wide, dependent_var = "BCTI_avg", independent_var = "reason_agreement")

# survey time
time_by_reason_model <- run_regression(exp4_wide %>% filter(duration_mins < 30), dependent_var = "duration_mins", independent_var = "reason_agreement")

# clicks
clicks_by_reason_model <- run_regression(exp4_wide, dependent_var = "sum_link_clicks", independent_var = "reason_agreement")

# extract descriptives for inline reporting
exp4_descriptives <- list(
  # Demographics
  n_subj = n_distinct(exp4_wide$id),
  n_subject_by_outcome = exp4_wide %>% 
    summarize(across(c("wgm_sciencegeneral", "reason_agreement", "BCTI_avg"), 
                     ~sum(!is.na(.x)
                     )
    )
    ),
  gender = exp4_wide %>% group_by(gender) %>% summarize(n = n_distinct(id)) %>% split(.$gender),
  age = exp4_wide %>% summarize(across(age,list(mean = mean, median = median, sd = sd), 
                                       .names = "{.fn}")) %>% rounded_numbers(),
  # RQ1 & RQ2
  means = exp4_wide %>% 
    summarize(across(c(avg_knowledge, avg_acceptance, avg_acceptance_initial), 
                     list(mean = mean, sd = sd), 
                     .names = "{.col}_{.fn}")) %>%
    rename_with(~sub("^avg_", "", .), everything()) %>% 
    mutate_if(is.numeric, round, digits = 2) %>% 
    mutate(across(ends_with("_mean"), ~paste0(.*100, " %"))) ,
  means_numeric = exp4_wide %>% 
    summarize(across(c(avg_knowledge, avg_acceptance, avg_acceptance_initial), 
                     list(mean = mean, sd = sd), 
                     .names = "{.col}_{.fn}")) %>%
    rename_with(~sub("^avg_", "", .), everything()),
  # H1
  # a
  cor_trust_knowledge = cor_trust_knowledge,
  # b (conditional on false responses)
  false_answers_cor_trust_acceptance = false_answers_cor_trust_acceptance,
  # b (all cases)
  cor_trust_acceptance = cor_trust_acceptance,
  # H2
  # a
  cor_conspiracy_knowledge = cor_conspiracy_knowledge,
  # b (conditional on false responses)
  false_answers_cor_conspiracy_acceptance = false_answers_cor_conspiracy_acceptance,
  # b(all cases)
  cor_conspiracy_acceptance = cor_conspiracy_acceptance, 
  # Exploratory
  # acceptance and knowledge
  conditional_acceptance = exp4_long %>% 
    pivot_longer(c(acceptance, acceptance_initial), 
                 names_to = "measure", 
                 values_to = "acceptance") %>% 
    group_by(knowledge, measure, acceptance) %>% 
    count() %>% 
    group_by(knowledge, measure) %>% 
    mutate (share = n/sum(n), 
            # rename knowledge values
            knowledge = ifelse(knowledge == FALSE, "false", "true")) %>% 
    mutate_if(is.numeric, round, digits = 3) %>% 
    mutate (share = paste0(share*100, " %")) %>% 
    super_split(knowledge,  measure, acceptance),
  # Justifications
  justifications_n = nrow(exp4_justifications), 
  justifications_n_participants = n_distinct(exp4_justifications$id),
  justification_by_category = exp4_justifications %>%
    group_by(category)%>%
    summarize(n = n(), 
              n_subjects = n_distinct(id)) %>%
    mutate(share = n/sum(n)) %>%
    mutate_if(is.numeric, round, digits=3) %>% 
    mutate(share = paste0(share*100, "%")) %>% 
    split(.$category),
  # Consensus acceptance
  acceptance_n = exp4_wide %>% 
    filter(!is.na(reason_agreement)) %>% 
    nrow(.),
  acceptance_by_reason = exp4_wide %>% 
    group_by(reason_agreement) %>% 
    summarize(n = n()) %>%
    drop_na(reason_agreement) %>% 
    mutate(share = n/sum(n)) %>%
    mutate_if(is.numeric, round, digits=3) %>% 
    mutate(share = paste0(share*100, "%")) %>% 
    split(.$reason_agreement),
  reason_followup_n = exp4_wide %>% 
    filter(!is.na(reason_followup)) %>% 
    nrow(.),
  # Exploratory: Differences by reason for trust
  reason_means = exp4_wide %>% 
    group_by(reason_agreement) %>% 
    summarize(across(c(wgm_sciencegeneral, avg_acceptance, sum_link_clicks, BCTI_avg), 
                     list(mean = mean, sd = sd), 
                     .names = "{.col}_{.fn}")) %>%
    rename_with(~sub("^avg_", "", .), everything()) %>% 
    mutate_if(is.numeric, round, digits = 2) %>% 
    mutate(across(acceptance_mean, ~paste0(.*100, " %"))) %>% 
    split(.$reason_agreement), 
  reason_trust_model = trust_by_reason_model %>% split(.$term), 
  reason_acceptance_model = acceptance_by_reason_model %>% split(.$term), 
  reason_clicks_model = clicks_by_reason_model %>% split(.$term),
  reason_BCTI_model = BCTI_by_reason_model %>%  split(.$term),
  reason_time_model = time_by_reason_model %>%  split(.$term), 
  n_participants_over30mins = exp4_wide %>% 
    filter(duration_mins >= 30) %>% 
    nrow(.)
)
```

```{r combined-data}
# combine data of all three studies
combined_data <- bind_rows(exp1_long %>% 
                             mutate(study = "1"), 
                           exp2_long %>% 
                             mutate(study = "2"), 
                           exp3_long %>% 
                             mutate(study = "3"),
                           exp4_long %>% 
                             mutate(study = "4")
                           ) %>% 
  # make a unique id variable
  mutate(id = paste0("study", study, "_", id))

# add a binned version for Conspiracy Thinking

# Define breaks 
breaks <- seq(0, 100, by = 20)
# Define labels for each bin in the desired format
labels <- paste0(c(0, 20, 40, 60, 80), "-", c(20, 40, 60, 80, 100))

combined_data <- combined_data %>% 
    # Create the binned version of CMQ_avg
  mutate(CMQ_avg_binned = cut(CMQ_avg, breaks = breaks, labels = labels, 
                              include.lowest = TRUE, right = FALSE))

# extract descriptives for inline reporting
combined_descriptives <- list(
  # acceptance
  mean_acceptance = combined_data %>% 
  group_by(acceptance) %>% 
  count() %>% 
    ungroup() %>% 
    mutate(share = n/sum(n)) %>% 
    mutate_if(is.numeric, round, digits = 3) %>% 
    mutate(share = paste0(share*100, " %")) %>% 
    split(.$acceptance), 
  # acceptance by trust
  acceptance_by_trust = combined_data %>% 
    group_by(acceptance, wgm_sciencegeneral) %>% 
    count() %>% 
    group_by(wgm_sciencegeneral) %>% 
    mutate(share = n/sum(n)) %>% 
    mutate_if(is.numeric, round, digits = 3) %>% 
    mutate(share = paste0(share*100, " %")) %>% 
    super_split(wgm_sciencegeneral, acceptance), 
  participants_by_trust = combined_data %>% 
  group_by(wgm_sciencegeneral) %>% 
  summarize(n_participants = n_distinct(id)) %>% 
    mutate(share = n_participants/sum(n_participants)) %>% 
    mutate_if(is.numeric, round, digits = 3) %>% 
    mutate(share = paste0(share*100, " %")) %>% 
    split(.$wgm_sciencegeneral),
  # acceptance by conspiracy thinking
  acceptance_by_conspiracy_thinking = combined_data %>% 
    group_by(acceptance, CMQ_avg_binned) %>% 
    count() %>% 
    group_by(CMQ_avg_binned) %>% 
    mutate(share = n/sum(n)) %>% 
    mutate_if(is.numeric, round, digits = 3) %>% 
    mutate(share = paste0(share*100, " %")) %>% 
    super_split(CMQ_avg_binned, acceptance),
  participants_by_conspiracy_thinking  = combined_data %>% 
    group_by(CMQ_avg_binned) %>% 
    summarize(n_participants = n_distinct(id)) %>% 
    mutate(share = n_participants/sum(n_participants)) %>% 
    mutate_if(is.numeric, round, digits = 3) %>% 
    mutate(share = paste0(share*100, " %")) %>% 
    split(.$CMQ_avg_binned),
  # acceptance by conspiracy belief
  acceptance_by_conspiracy_belief = combined_data %>% 
    # bring to long format
    pivot_longer(cols = BCTI_apollo:BCTI_evolution, 
                 names_to = "BCTI_item", values_to = "score") %>% 
    mutate(BCTI_item = sub("^BCTI_", "", BCTI_item)) %>% 
    group_by(acceptance, BCTI_item, score) %>% 
    count() %>% 
    group_by(BCTI_item, score) %>% 
    mutate(share = n/sum(n)) %>% 
    mutate_if(is.numeric, round, digits = 3) %>% 
    mutate(share = paste0(share*100, " %")) %>% 
    super_split(acceptance, BCTI_item, score),
  participants_by_conspiracy_belief  = combined_data %>% 
    # bring to long format
    pivot_longer(cols = BCTI_apollo:BCTI_evolution, 
                 names_to = "BCTI_item", values_to = "score") %>% 
    mutate(BCTI_item = sub("^BCTI_", "", BCTI_item)) %>% 
    group_by(BCTI_item, score) %>% 
    summarize(n_participants = n_distinct(id)) %>% 
    mutate(share = n_participants/sum(n_participants)) %>% 
    mutate_if(is.numeric, round, digits = 3) %>% 
    mutate(share = paste0(share*100, " %")) %>% 
    super_split(BCTI_item, score),
  # knowledge
  mean_knowledge = combined_data %>% 
    group_by(knowledge) %>% 
    count() %>% 
    ungroup() %>% 
    mutate (share = n/sum(n), 
            # rename knowledge values
            knowledge = ifelse(knowledge == FALSE, "false", "true")) %>% 
    mutate_if(is.numeric, round, digits = 3) %>% 
    mutate(share = paste0(share*100, " %")) %>% 
    split(.$knowledge),
  # flat earthers with no trust at all in science
  # number of participants
  flat_earth_no_trust_nparticipants = combined_data %>% 
    filter(BCTI_flatearth == 9 & wgm_sciencegeneral == 1) %>%
    summarize(n_participants = n_distinct(id)) %>% 
    pull(),
  # acceptance
  flat_earth_no_trust_acceptance = combined_data %>% 
    filter(BCTI_flatearth == 9 & wgm_sciencegeneral == 1) %>%
    group_by(acceptance) %>% 
    summarise(n = n()) %>% 
    mutate(share = n/sum(n)) %>% 
    mutate_if(is.numeric, round, digits = 3) %>% 
    mutate(share = paste0(share*100, " %")) %>% 
    split(.$acceptance), 
  # Lowest decile acceptance
  lowest_decile_acceptance = combined_data %>% 
    group_by(id) %>% 
    summarise(acceptance_rate = sum(acceptance == "Yes")/n(), 
              n = n()) %>% 
    arrange(acceptance_rate) %>% 
    slice(1:(n() * 0.1)) %>%  # Get the lowest 10% (decile)
    summarize(mean_acceptance_rate = mean(acceptance_rate, na.rm = TRUE), 
              n_smaller_or_equalto_0.5 = sum(acceptance_rate <=0.5)) %>% 
    mutate_if(is.numeric, round, digits = 3) %>% 
    mutate(mean_acceptance_rate = paste0(mean_acceptance_rate*100, " %")), 
  # Lowest decile knowledge
  lowest_decile_knowledge = combined_data %>% 
    group_by(id) %>% 
    summarise(acceptance_rate = sum(acceptance == "Yes")/n(), 
              knowledge_rate = mean(knowledge, na.rm=TRUE),
              n = n()) %>% 
    arrange(knowledge_rate) %>% 
    slice(1:(n() * 0.1)) %>%  # Get the lowest 10% (decile)
    summarize(mean_acceptance_rate = mean(acceptance_rate, na.rm = TRUE), 
              mean_knowledge_rate = mean(knowledge_rate, na.rm = TRUE)) %>% 
    mutate_if(is.numeric, round, digits = 3) %>% 
    mutate(mean_acceptance_rate = paste0(mean_acceptance_rate*100, " %"), 
           mean_knowledge_rate = paste0(mean_knowledge_rate*100, " %"))
)
```

## Deviations from preregistration

For Study 2, we restricted our main hypotheses about acceptance to cases in which participants initially provided a wrong answer. However, this meant the more participants had initially provided correct answers, the fewer opportunities they had for accepting correct answers. We provide results on these conditional correlations--for Study 2 and for all other studies--in the ESM, section \@ref(exp1) to \@ref(exp4) [^1]. However, for the analysis presented here, we proceeded as preregistered for all other studies, by reporting unconditional correlations between acceptance and trust in science, or, respectively, conspiracy belief.

[^1]: Only in Study 4 do we find evidence that changing one's mind towards the scientific consensus is associated with (more) trust in science (Studies 1: r = `r exp1_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp1_descriptives$false_answers_cor_trust_acceptance$p.value`; 2: r = `r exp2_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp2_descriptives$false_answers_cor_trust_acceptance$p.value`; 3: r = `r exp3_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp3_descriptives$false_answers_cor_trust_acceptance$p.value`; 4: r = `r exp4_descriptives$false_answers_cor_trust_acceptance$estimate`, p `r exp4_descriptives$false_answers_cor_trust_acceptance$p.value`) and only in Study 2 evidence that it is associated with (less) conspiracy beliefs and (less) conspiracy thinking (Studies 1: r = `r exp1_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp1_descriptives$false_answers_cor_conspiracy_acceptance$p.value`; 2: r = `r exp2_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp2_descriptives$false_answers_cor_conspiracy_acceptance$p.value`; 3: r = `r exp3_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp3_descriptives$false_answers_cor_conspiracy_acceptance$p.value`; 4: r = `r exp4_descriptives$false_answers_cor_conspiracy_acceptance$estimate`, p `r exp4_descriptives$false_answers_cor_conspiracy_acceptance$p.value`).

## Procedure

After providing their consent to participate in the study, participants were given an attention check "While watching the television, have you ever had a fatal heart attack?" [1-6; 1 = Never, 6 = Often]. All participants who did not answer "1 = Never" were excluded. Participants then read the following instructions:"We will ask you 10 questions about science. After each question, we will provide you with the scientifically consensual answer and ask whether you accept it." Next, participants answered a set of 10 basic science questions in random order. After each question, participants were presented with an answer reflecting the scientific consensus, and asked whether they accepted it. In Studies 2 and 3, participants additionally saw a short explanation, partly based on explanations generated by ChatGPT, and three links to authoritative sources supporting the answer. In Study 4, we provided only two links and no explanation. Participants then answered questions on conspiracy thinking, conspiracy beliefs, and trust in science.

In Studies 2, 3, and 4, we presented participants with open-ended questions so they could explain their rejection of the scientific consensus. In Studies 3 and 4, we additionally gave participants the option to change their answer and accept the scientific consensus. Finally, at the end of Studies 3 and 4, we asked participants: "For the questions in which you agreed with the scientific consensus, would you say that...?" The answer options were: (i) "You mostly agree with the consensus because, on that question, you trust scientists", (ii) "You mostly agree with the consensus because you have been able to independently verify it", and (iii) "Other", with a text box for participants to explain. Participants who selected “You mostly agree with the consensus because you have been able to independently verify it”, were asked the open-ended follow-up question: "Could you please tell us how you independently verified the information?".

## Participants

After removing failed attention checks, the total sample size was `r exp1_descriptives$n_subj + exp2_descriptives$n_subj + exp3_descriptives$n_subj + exp4_descriptives$n_subj` (`r exp1_descriptives$n_subj` in Study 1, six failed attention checks; `r exp2_descriptives$n_subj` in Study 2, 11 failed attention checks; `r exp3_descriptives$n_subj` in Study 3, no failed attention checks; `r exp4_descriptives$n_subj` in Study 4, two failed attention checks) participants in the US, recruited through Prolific. Details and demographics can be found in the online supplemental material. While samples for Studies 1 and 2 were convenience samples, Studies 3 and 4 were conducted on a sample holding vaccine-skeptic beliefs. Prolific allows selecting participants based on their answers to a range of questions. We picked three of these questions and only recruited participants who met our criteria for each of them:

1.  "Please describe your attitudes towards the COVID-19 (Coronavirus) vaccines: [For (I feel positively about the vaccines); Against (I feel negatively about the vaccines); Neutral (I don't have strong opinions either way); Prefer not to say]". We selected participants who answered "Against".
2.  "Have you received a coronavirus (COVID-19) vaccination? [Yes (at least one dose); No; Prefer not to answer]". We select only people who answered "No".
3.  "On a scale from 1-7, please rate to what extent you agree with the following statement: I believe that scheduled immunizations are safe for children. [1 (totally disagree); 2 (disagree); 3 (somewhat disagree); 4 (neither agree nor disagree); 5 (somewhat agree); 6 (agree); 7 (totally agree); rather not say]". We select only people who answered "1", "2", or "3".

## Materials

### Scientific facts

Studies 1 to 3 used 10 facts drawn from widely used questionnaires about science knowledge [@allumScienceKnowledgeAttitudes2008; @durantPublicUnderstandingScience1989; @millerMeasurementCivicScientific1998a] sometimes referred to as the "Oxford scale" [@gauchatCulturalAuthorityScience2011]. A ‘trick’ question was added in Study 1 and removed as its wording proved unclear. Study 4 used 10 more recent scientific discoveries. Table \@ref(tab:knowledge) shows all questions and their answer options.

```{r knowledge}
# Function to replace special characters in all cells of a data frame
replace_special_characters <- function(df) {
  df %>%
    mutate(across(everything(), ~ gsub("\u2013|\u2014|\u2212", "-", .))) %>% # Replace en-dash, em-dash, and Unicode minus with ASCII hyphen-minus
    mutate(across(everything(), ~ iconv(., from = "UTF-8", to = "ASCII//TRANSLIT"))) # Ensure all text is ASCII
}

# Read the CSV file
items <- read_csv("exp_4/materials/overview_questions.csv")

# Replace special characters
items_cleaned <- replace_special_characters(items) %>% 
  mutate(id = 1:nrow(.)) %>%
  select(id, everything())  # Ensure 'id' is the first column

# add footnote for the tree items
items_cleaned[11, 2] <- paste0("*", 
                               items_cleaned[11, 2])

# Output the table
knitr::kable(items_cleaned, booktabs = T, longtable = TRUE,
    caption = "Science knowledge items", 
    full_width = T, 
    col.names = c("", "Study 1-3", "Study 4")) %>%
  kable_styling(font_size = 8) %>%
  column_spec(1, width = "2em") %>%
  column_spec(2, width = "22em")%>%
  column_spec(3, width = "22em") %>% 
  footnote(symbol = "Only used in Study 1")
```

### Conspiracy beliefs

We selected 10 science/health related conspiracy theories from the Belief in Conspiracy Theory Inventory (BCTI) [@pennycookOverconfidentlyConspiratorialConspiracy2022] (Table \@ref(tab:conspiracy)). Participants were asked: "Below is a list of events for which the official version has been disputed. For each event, we would like you to indicate to what extent you believe the cover-up version of events is true or false. [1-9; labels: 1 - completely false, 5 - unsure, 9 - completely true]".

### Conspiracy thinking

For all results presented here, we used the four-item conspiracy mentality questionnaire (CMQ) [@bruderMeasuringIndividualDifferences2013]. We also assessed the single item conspiracy beliefs scale (SICBS) [@lantianMeasuringBeliefConspiracy2016]. Details and comparisons between the scales can be found in the ESM, sections \@ref(exp1) to \@ref(exp4).

```{r conspiracy, echo=FALSE}
# Create the data frame
items <- c(
  "The Apollo moon landings never happened and were staged in a Hollywood film studio.",
  "A cure for cancer was discovered years ago, but this has been suppressed by the pharmaceutical industry and the U.S. Food and Drug Administration (FDA).",
  "The spread of certain viruses and/or diseases is the result of the deliberate, concealed efforts of vested interests.",
  "The claim that the climate is changing due to emissions from fossil fuels is a hoax perpetrated by corrupt scientists who want to spend more taxpayer money on climate research.",
  "The Earth is flat (not spherical) and this fact has been covered up by scientists and vested interests.",
  "There is a causal link between vaccination and autism that has been covered up by the pharmaceutical industry.",
  "In the 1950s and 1960s more than 100 million Americans received a polio vaccine contaminated with a potentially cancer-causing virus.",
  "Proof of alien contact is being concealed from the public.",
  "Hydroxychloroquine has been demonstrated to be a safe and effective treatment of COVID and this information is being suppressed.",
  "Dinosaurs never existed, evolution is not real, and scientists have been faking the fossil record.")

conspiracy_items <- data.frame(id = 1:length(items), items = items)

# Output the table
kbl(conspiracy_items, booktabs = T, longtable = T, col.names = NULL,
    caption = "Conspiracy items", full_width = T) %>%
  kable_styling(font_size = 8) %>% 
  column_spec(1, width = "3em") %>%
  column_spec(2, width = "40em") %>% 
  footnote(general = "Participants were asked to rate their belief in the conspiracy on a scale from 1 to 9, with the labels: `1 - completely false, 5 - unsure, 9 - completely true`.",
           escape = FALSE,
           footnote_as_chunk = TRUE,
           threeparttable = TRUE)
```

### Trust in science

In all analyses reported in the main paper, we measure trust in science via a question selected from the Wellcome Global Monitor surveys [@wellcomeglobalmonitorWellcomeGlobalMonitor2018; @wellcomeglobalmonitorWellcomeGlobalMonitor2020]: "In general, would you say that you trust science a lot, some, not much, or not at all? [1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]". We chose this question as it seemed to be the most general one. In the ESM (sections \@ref(exp1) to \@ref(exp4)), we additionally report results for two alternative measures of trust included in our studies: Another from the WGM surveys ("How much do you trust scientists in this country? Do you trust them a lot, some, not much, or not at all? [1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]"), and one from the Pew Research Center [e.g. @funkKeyFindingsPublic2019] ("How much confidence do you have in scientists to act in the best interests of the public? [1-5; 1 = No confidence at all, 5 = A great deal of confidence]"), the latter having been used in a recent international study on trust in science [@colognaTrustScientistsTheir2024]. We selected these items so that we could compare the answers in our sample to global survey results. We find that all three items are generally highly correlated throughout all studies, and that our results reported here generally replicate when using either of the alternatives measures (with two exceptions: In Study 3, we find no correlation between acceptance and the Pew question; In Study 4, we find a correlation between knowledge and both alternative trust measures, but not with our main measure; see ESM, sections \@ref(exp3) and \@ref(exp4)).

# Results

(ref:summary-plot) Points represent the average share of acceptance and numbers the absolute count of participants as a function of: **A** the level of trust in science ("In general, would you say that you trust science a lot, some, not much, or not at all? [1 = Not at all, 2 = Not much, 3 = Some, 4 = A lot]"); **B** the average conspiracy thinking (CMQ, five items on a scale from 0 to 100); **C** the belief in specific conspiracy theories (i.e. participants who answered 9, "completely true", for a given theory, see Table \@ref(tab:conspiracy) for the list of the theories).

```{r summary-plot, fig.cap="(ref:summary-plot)", fig.height= 10, fig.width=10}
# make plot data for trust
plot_trust <- combined_data %>% 
  drop_na(wgm_sciencegeneral, acceptance) %>% 
  # make a labels version of trust in science
  mutate(wgm_sciencegeneral = case_when(
    wgm_sciencegeneral == 1 ~ "1 (Not at all)",
    wgm_sciencegeneral == 2 ~ "2 (Not much)",
    wgm_sciencegeneral == 3 ~ "3 (Some)",
    wgm_sciencegeneral == 4 ~ "4 (A lot)",
    TRUE ~ as.character(wgm_sciencegeneral)  # Handle any other cases (optional)
  )) %>% 
  group_by(study, wgm_sciencegeneral, acceptance) %>% 
  summarize(n = n(), 
            n_participants = n_distinct(id)) %>% 
  group_by(study, wgm_sciencegeneral) %>% 
  mutate (share = n/sum(n), 
          n_participants = max(n_participants)) %>% 
  mutate_if(is.numeric, round, digits = 3)

# make trust plot
trust <- ggplot(plot_trust %>% 
                  filter(acceptance == "Yes"), 
                aes(x = wgm_sciencegeneral, y = share, color = study, shape = study)) +
  geom_pointrange(aes(ymin = 0, ymax = share), 
                  position = position_dodge(width = 0.8), 
                  alpha = 0.5, 
                  show.legend = FALSE) +
  geom_text(aes(label = n_participants, y = share + 0.01),  # Align text labels with dodged points
            position = position_dodge(width = 0.8), 
            vjust = -0.5, 
            size = 3, 
            show.legend = FALSE) +
  #geom_point(position = "dodge", alpha = 0.5) +
  # geom_text_repel(aes(label = paste0(n_participants)),
  #                 vjust = -0.5, size = 3,
  #                 show.legend = FALSE) +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1), limits = c(0, 1.02)) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  labs(x = "Trust in science", y = "Acceptance rate of consensus", color = "Study", shape = "Study") +
  plot_theme + 
  theme(legend.position = "top")

# make plot data for conspiracy belief
plot_conspiracy <- combined_data %>% 
  # bring to long format
  pivot_longer(cols = BCTI_apollo:BCTI_evolution, 
               names_to = "BCTI_item", values_to = "score") %>% 
  # make a belief variable if participant indicates a numeric answer above the
  # a belief threshold 
  mutate(belief = ifelse(score == 9, TRUE, 
                         ifelse(is.na(score), NA, FALSE)
                         ),
         BCTI_item = sub("^BCTI_", "", BCTI_item),
         # rename BCTI items
         BCTI_item = case_when(
           BCTI_item == "apollo" ~ "Moon landing", 
           BCTI_item == "cancer" ~ "Cancer cure", 
           BCTI_item == "viruses" ~ "Viruses",
           BCTI_item == "climatechange" ~ "Climate change",
           BCTI_item == "flatearth" ~ "Flat earth",
           BCTI_item == "autism" ~ "Vaccine autism",
           BCTI_item == "polio" ~ "Vaccine cancer",
           BCTI_item == "alien" ~ "Aliens",
           BCTI_item == "hydorxychlor" ~ "Covid hydroxychloroquine",
           BCTI_item == "evolution" ~ "Evolution"
         )
         ) %>% 
  group_by(study, BCTI_item, belief, acceptance) %>% 
  summarize(n = n(),
            n_participants = n_distinct(id)) %>% 
  group_by(study, BCTI_item, belief) %>% 
  mutate (share = n/sum(n), 
          n_participants = max(n_participants)) %>% 
  mutate_if(is.numeric, round, digits = 3) %>% 
  ungroup()

# make conspiracy belief plot
conspiracy <- ggplot(plot_conspiracy %>% 
         filter(acceptance == "Yes" & belief == TRUE), 
         aes(x = BCTI_item, y = share, color = study, shape = study)) +
  geom_pointrange(aes(ymin = 0, ymax = share), 
                  position = position_dodge(width = 0.8), 
                  alpha = 0.5, 
                  show.legend = FALSE) +
  geom_text(aes(label = n_participants, y = share + 0.01),  # Align text labels with dodged points
            position = position_dodge(width = 0.8), 
            vjust = -0.5, 
            size = 3, 
            show.legend = FALSE) +
  # geom_point(position = "dodge", alpha = 0.5) +
  # geom_text_repel(aes(label = paste0(n_participants)),
  #                 vjust = -0.5, size = 3,
  #                 show.legend = FALSE) +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1), limits = c(0, 1.02)) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  labs(x = "Conspiracy theory belief", y = "Acceptance rate of consensus", color = "Study", shape = "Study") +
  plot_theme +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

# Conspiracy Thinking

# Define breaks for each integer from 1 to 9
breaks <- seq(0, 100, by = 20)

# Define labels for each bin in the desired format
labels <- paste0(c(0, 20, 40, 60, 80), "-", c(20, 40, 60, 80, 100))

plot_data <- combined_data %>% 
  drop_na(CMQ_avg) %>% 
  # Create the binned version of CMQ_avg
  mutate(CMQ_avg_binned = cut(CMQ_avg, breaks = breaks, labels = labels, include.lowest = TRUE, right = FALSE)) %>% 
  group_by(study, CMQ_avg_binned, acceptance) %>% 
  summarize(n = n()) %>% 
  group_by(study, CMQ_avg_binned) %>% 
  mutate (share = n/sum(n)) %>% 
  mutate_if(is.numeric, round, digits = 3)

n_participants <- combined_data %>% 
  drop_na(CMQ_avg) %>% 
  # Create the binned version of CMQ_avg
  mutate(CMQ_avg_binned = cut(CMQ_avg, breaks = breaks, labels = labels, include.lowest = TRUE, right = FALSE)) %>% 
  group_by(study, CMQ_avg_binned) %>% 
  summarize(n_participants = n_distinct(id))

plot_conspiracy_thinking <- left_join(plot_data, n_participants)

# make plot
conspiracy_thinking  <- ggplot(plot_conspiracy_thinking  %>% 
                                 filter(acceptance == "Yes"), 
                               aes(x = CMQ_avg_binned, y = share, color = study, shape = study)) +
  geom_pointrange(aes(ymin = 0, ymax = share), 
                  position = position_dodge(width = 0.8), 
                  alpha = 0.5) +
  geom_text(aes(label = n_participants, y = share + 0.01),  # Align text labels with dodged points
            position = position_dodge(width = 0.8), 
            vjust = -0.5, 
            size = 3, 
            show.legend = FALSE) +
  # geom_point(position = "dodge", alpha = 0.5) +
  # geom_text_repel(aes(label = paste0(n_participants)),
  #                 vjust = -0.5, size = 3,
  #                 show.legend = FALSE) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1), limits = c(0, 1.02)) +
  labs(x = "Conspiracy Thinking", y = "Acceptance rate of consensus", color = "Study", shape = "Study") +
  plot_theme + 
  theme(legend.position = "top")

# remove elements of plots for better overall plot
conspiracy_thinking <- conspiracy_thinking + 
  #rremove("ylab") +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) 
  #guides(color = "none", shape = "none")

conspiracy <- conspiracy +
  rremove("ylab") +
  guides(color = "none", shape = "none")

trust <- trust +
  rremove("ylab")

# common plot
patchwork <- (trust + conspiracy_thinking)/conspiracy + 
  plot_layout(guides = 'collect', widths = c(1, 1)) +
  plot_annotation(tag_levels = 'A')  & theme(legend.position = "top")

wrap_elements(patchwork) +
  labs(tag = "Rate of acceptance of consensual scientific facts") +
    theme(
    plot.tag = element_text(angle = 90, , face = "bold", size = 12),
    plot.tag.position = "left"
  )
```

The main outcome of interest is acceptance of the scientifically consensual facts presented. Overall, acceptance was very high (aggregating across all studies: `r combined_descriptives$mean_acceptance$Yes$share`; Studies 1: `r exp1_descriptives$means$acceptance_mean`; 2: `r exp2_descriptives$means$acceptance_mean`; 3: `r exp3_descriptives$means$acceptance_mean`; 4: `r exp4_descriptives$means$acceptance_mean`). Note that this includes both participants who had previously correctly answered the knowledge question, and participants who changed their mind when presented with the scientific consensus. In Studies 3 and 4, we gave participants a second chance in case they had initially rejected the consensus, which slightly increased acceptance rates in those studies (initial acceptance in Studies 3: `r exp3_descriptives$means$acceptance_initial_mean`; 4: `r exp4_descriptives$means$acceptance_initial_mean`).

As shown in Figure \@ref(fig:summary-plot), these very high rates of acceptance hold for: participants who do not trust science at all (`r combined_descriptives$participants_by_trust$'1'$share` of participants, acceptance rate of `r combined_descriptives$acceptance_by_trust$'1'$Yes$share`), participants who rank in the top two deciles of the conspiracy thinking scale (`r combined_descriptives$participants_by_conspiracy_thinking$'80-100'$share` of participants, acceptance rate of `r combined_descriptives$acceptance_by_conspiracy_thinking$'80-100'$Yes$share`), participants who consider as "completely true" (the maximum of the 9-point scale) conspiracy theories stating that the earth is flat (`r combined_descriptives$participants_by_conspiracy_belief$flatearth$'9'$share` of participants, acceptance rate of `r combined_descriptives$acceptance_by_conspiracy_belief$Yes$flatearth$'9'$share`), or that climate change due to fossil emissions is a hoax (`r combined_descriptives$participants_by_conspiracy_belief$climatechange$'9'$share` of participants, acceptance rate of `r combined_descriptives$acceptance_by_conspiracy_belief$Yes$climatechange$'9'$share`).

```{r}
#combined_descriptives$flat_earth_no_trust_nparticipants
```

Participants in the lowest decile of acceptance still had an average acceptance rate of `r combined_descriptives$lowest_decile_acceptance$mean_acceptance_rate`. Even the three participants who considered as "completely true" that the earth is flat and who said they do "not trust science at all" had an average acceptance rate of `r combined_descriptives$flat_earth_no_trust_acceptance$Yes$share`.

These high acceptance rates do not merely reflect science knowledge: participants only correctly answered `r combined_descriptives$mean_knowledge$true$share` of the questions (Studies 1: `r exp1_descriptives$means$knowledge_mean`; 2: `r exp2_descriptives$means$knowledge_mean`; 3: `r exp3_descriptives$means$knowledge_mean`; 4: `r exp4_descriptives$means$knowledge_mean`) before they were provided with the scientifically consensual answer. Even the lowest decile in science knowledge, which on average answered correctly only on `r combined_descriptives$lowest_decile_knowledge$mean_knowledge_rate` of the questions, had an average acceptance rate of `r combined_descriptives$lowest_decile_knowledge$mean_acceptance_rate`.

Did participants who had initially provided a wrong answer change their minds towards the scientific consensus? Yes. In most cases (Studies 1: `r exp1_descriptives$conditional_acceptance$false$Yes$share`; 2: `r exp2_descriptives$conditional_acceptance$false$Yes$share`; 3: `r exp3_descriptives$conditional_acceptance$false$acceptance$Yes$share`; 4: `r exp4_descriptives$conditional_acceptance$false$acceptance$Yes$share`), participants readily accepted the scientific consensus after having initially given the wrong answer to a question.

Regarding H1, we find a consistent association between trust in science and acceptance of the scientific consensus (Studies 1: r = `r exp1_descriptives$cor_trust_acceptance$estimate`, p `r exp1_descriptives$cor_trust_acceptance$p.value`; 2: r = `r exp2_descriptives$cor_trust_acceptance$estimate`, p `r exp2_descriptives$cor_trust_acceptance$p.value`; 3: r = `r exp3_descriptives$cor_trust_acceptance$estimate`, p `r exp3_descriptives$cor_trust_acceptance$p.value`; 4: r = `r exp4_descriptives$cor_trust_acceptance$estimate`, p `r exp4_descriptives$cor_trust_acceptance$p.value`), but a less consistent relation between trust in science and science knowledge (Studies 1: r = `r exp1_descriptives$cor_trust_knowledge$estimate`, p `r exp1_descriptives$cor_trust_knowledge$p.value`; 2: r = `r exp2_descriptives$cor_trust_knowledge$estimate`, p `r exp2_descriptives$cor_trust_knowledge$p.value`; 3: r = `r exp3_descriptives$cor_trust_knowledge$estimate`, p `r exp3_descriptives$cor_trust_knowledge$p.value`; 4: r = `r exp4_descriptives$cor_trust_knowledge$estimate`, p `r exp4_descriptives$cor_trust_knowledge$p.value`).

Regarding H2, the results are mixed for the relation of conspiracy beliefs (measured as the average acceptance of all the conspiracy beliefs) with both acceptance of the scientific consensus (Studies 1: r = `r exp1_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp1_descriptives$cor_conspiracy_acceptance$p.value`; 2: r = `r exp2_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp2_descriptives$cor_conspiracy_acceptance$p.value`; 3: r = `r exp3_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp3_descriptives$cor_conspiracy_acceptance$p.value`; 4: r = `r exp4_descriptives$cor_conspiracy_acceptance$estimate`, p `r exp4_descriptives$cor_conspiracy_acceptance$p.value`) and science knowledge (Studies 1: r = `r exp1_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp1_descriptives$cor_conspiracy_knowledge$p.value`; 2: r = `r exp2_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp2_descriptives$cor_conspiracy_knowledge$p.value`; 3: r = `r exp3_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp3_descriptives$cor_conspiracy_knowledge$p.value`; 4: r = `r exp4_descriptives$cor_conspiracy_knowledge$estimate`, p `r exp4_descriptives$cor_conspiracy_knowledge$p.value`).

Why did participants reject the scientific consensus? We collected a total of `r exp4_descriptives$justifications_n + exp3_descriptives$justifications_n + exp2_descriptives$justifications_n` answers (Studies 2: `r exp2_descriptives$justifications_n`; 3: `r exp3_descriptives$justifications_n`; 4: `r exp4_descriptives$justifications_n`) from `r exp4_descriptives$justifications_n_participants + exp3_descriptives$justifications_n_participants + exp2_descriptives$justifications_n_participants` (Studies 2: `r exp2_descriptives$justifications_n_participants`; 3: `r exp3_descriptives$justifications_n_participants`; 4: `r exp4_descriptives$justifications_n_participants`) participants to the open-ended questions on why they had rejected the scientific consensus on a particular question. Based on the answers, we created five categories (Table \@ref(tab:justifications)). All answers can be read in the ESM, sections \@ref(exp2) to \@ref(exp4).

```{r justifications}
## make a common table of exp 2 and exp 3
justifications <- bind_rows(exp2_justifications %>% 
                              mutate(study = "study 2"), 
                            exp3_justifications %>% 
                              mutate(study = "study 3", 
                                     final_coding = as.character(final_coding)),
                            exp4_justifications %>% 
                              mutate(study = "study 4"))

table <- justifications %>%
  group_by(category)%>%
  summarize(n = n(), 
            n_subjects = n_distinct(id)) %>%
  mutate(Share = n/sum(n)) %>%
  mutate_if(is.numeric, round, digits=3) %>% 
  mutate(Share = paste0(Share*100, "%")) %>% 
  arrange(desc(n)) %>% 
  select(category, n, Share, n_subjects) %>% 
  ungroup() %>% 
  rename(
    Category = category,
    `N (instances)` = n, 
    `Share (instances)` = Share, 
    `N (participans)*` = n_subjects
  ) %>% 
  # remove decimal places
  mutate_if(is.numeric, ~as.integer(.x))

table %>%
  apa_table(caption = "Justifications for rejecting the scientific consensus by category, Studies 2, 3 and 4 combined.", 
            note = "*Participans with at least one answer in that category")

# table %>% kbl(
#   booktabs = T, longtable = T,
#     caption = "Justifications for rejecting the scientific consensus by category, Studies 2, 3 and 4 combined.", 
#   full_width = T) %>%
#   footnote(general = "*Participans with at least one answer in that category",
#            escape = FALSE,
#            footnote_as_chunk = TRUE,
#            threeparttable = TRUE)
```

Why did participants say they accept the scientific consensus? In Studies 3 and 4--the vaccine hesitant samples--we had asked participants about cases in which they agreed with the scientific consensus. A total of `r exp3_descriptives$acceptance_n + exp4_descriptives$acceptance_n` (Studies 3: `r exp3_descriptives$acceptance_n`; 4: `r exp4_descriptives$acceptance_n`) participants answered this question. There were more participants saying they accepted the scientific consensus because they independently verified the fact (Studies 3: `r exp3_descriptives$acceptance_by_reason[["independent verification"]]$share`; 4: `r exp4_descriptives$acceptance_by_reason[["independent verification"]]$share` ), than participants saying it was because they trust scientists (Studies 3: `r exp3_descriptives$acceptance_by_reason[["trust in scientists"]]$share`; 4: `r exp4_descriptives$acceptance_by_reason[["trust in scientists"]]$share`)[^2]. Answers to a question about how they had done so can be found in the ESM, sections \@ref(exp3) and \@ref(exp4).

[^2]: `r exp3_descriptives$acceptance_by_reason[["other"]]$share` in Study 3 and `r exp4_descriptives$acceptance_by_reason[["other"]]$share` in Study 4 answered with "other" and gave an open-ended explanation (see ESM, sections \@ref(exp3) and \@ref(exp4)).

```{r}
#exp4_descriptives$reason_time_model$'reason_agreementtrust in scientists'$estimate %>% round(digits = 0) %>% abs()
#median(combined_data$duration_mins) %>% round(digits = 0)
#exp3_descriptives$n_participants_over30mins
#exp4_descriptives$n_participants_over30mins
```

In an exploratory analysis, we ran linear regressions to test whether there are differences between participants who said they had trusted science and those who said they had verified the information independently. Participants who said they accepted the consensus because of trust in scientists reported trusting science more (Studies 3: mean = `r exp3_descriptives$reason_means$'trust in scientists'$wgm_sciencegeneral_mean`; $\hat{\beta}_{\text{Trust}}$ = `r exp3_descriptives$reason_trust_model$'reason_agreementtrust in scientists'$estimate`, p `r exp3_descriptives$reason_trust_model$'reason_agreementtrust in scientists'$p.value` on a scale from 1 to 4; 4: mean = `r exp4_descriptives$reason_means$'trust in scientists'$wgm_sciencegeneral_mean`; $\hat{\beta}_{\text{Trust}}$ = `r exp4_descriptives$reason_trust_model$'reason_agreementtrust in scientists'$estimate`, p `r exp4_descriptives$reason_trust_model$'reason_agreementtrust in scientists'$p.value`) than those who said they verified independently (Studies 3 mean = `r exp3_descriptives$reason_means$'independent verification'$wgm_sciencegeneral_mean`; 4: mean = `r exp4_descriptives$reason_means$'independent verification'$wgm_sciencegeneral_mean`). We did not find a difference regarding acceptance (Studies 3: $\hat{\beta}_{\text{Acceptance}}$ = `r exp3_descriptives$reason_acceptance_model$'reason_agreementtrust in scientists'$estimate`, p `r exp3_descriptives$reason_acceptance_model$'reason_agreementtrust in scientists'$p.value` on a scale from 0 to 1; 4: $\hat{\beta}_{\text{Acceptance}}$ = `r exp4_descriptives$reason_acceptance_model$'reason_agreementtrust in scientists'$estimate`, p `r exp4_descriptives$reason_acceptance_model$'reason_agreementtrust in scientists'$p.value`) or regarding beliefs in conspiracy theories (Studies 3: $\hat{\beta}_{\text{BCTI}}$ = `r exp3_descriptives$reason_BCTI_model$'reason_agreementtrust in scientists'$estimate`, p `r exp3_descriptives$reason_BCTI_model$'reason_agreementtrust in scientists'$p.value` on a scale from 1 to 9; 4: $\hat{\beta}_{\text{BCTI}}$ = `r exp4_descriptives$reason_BCTI_model$'reason_agreementtrust in scientists'$estimate`, p `r exp4_descriptives$reason_BCTI_model$'reason_agreementtrust in scientists'$p.value`). We also did not find a difference in time spent on the survey in Study 3 ($\hat{\beta}_{\text{Time}}$ = `r exp3_descriptives$reason_time_model$'reason_agreementtrust in scientists'$estimate`, p `r exp3_descriptives$reason_time_model$'reason_agreementtrust in scientists'$p.value`; median = `r median(exp3_wide$duration_mins)` mins), but in Study 4, people who said they had accepted the consensus because they they trust scientists tended to spend on average two minutes less on the survey ($\hat{\beta}_{\text{Time}}$ = `r exp4_descriptives$reason_time_model$'reason_agreementtrust in scientists'$estimate`, p `r exp4_descriptives$reason_time_model$'reason_agreementtrust in scientists'$p.value`; median = `r median(exp4_wide$duration_mins)` mins).[^3] In Study 4, in which we used facts that participants were unlikely to have encountered before, we tracked whether people clicked on the source links we provided--a behavior that you would expect from people who report verifying facts independently. On average, participants clicked only on `r exp4_descriptives$reason_means$'trust in scientists'$sum_link_clicks_mean` links (out of 20 possible clicks) and there was no difference between the two groups ($\hat{\beta}_{\text{Clicks}}$ = `r exp4_descriptives$reason_clicks_model$'reason_agreementtrust in scientists'$estimate`, p `r exp4_descriptives$reason_clicks_model$'reason_agreementtrust in scientists'$p.value`).

[^3]: For these analyses, we excluded outliers who took over 30 mins for the survey, which was estimated to take around 10 mins, and for which the median time was seven minutes. As a result, we excluded one participant in Study 3 and four in Study 4. Significance levels are not affected by these exclusions.

More detailed results addressing all our pre-registered research questions can be found in the ESM, sections \@ref(exp1) to \@ref(exp4).

# Discussion

US participants were asked whether they accepted scientifically consensual answers on basic science questions.We found quasi-universal acceptance of basic science, with an overall rate of acceptance of `r combined_descriptives$mean_acceptance$Yes$share`, which remained very high for participants who declared not trusting science at all (`r combined_descriptives$acceptance_by_trust$'1'$Yes$share` of acceptance), or who endorsed theories blatantly violating scientific knowledge, such as flat earth (`r combined_descriptives$acceptance_by_conspiracy_belief$Yes$flatearth$'9'$share` of acceptance).

Overall, the present findings support the motivated rejection of science model [@lewandowskyMotivatedRejectionScience2016], in which “people tend to reject findings that threaten their core beliefs or worldview” (p. 217). A number of participants in our studies rejected basic tenets of science (on evolution, the shape of the earth, etc.). However, they still accepted the vast majority of basic scientific knowledge presented to them. This suggests that these participants had reasons to reject specific scientific knowledge, and that this rejection prompted them to express a lower trust in science when asked general questions on the topic. These participants did not appear to have general grounds for distrusting science, which should have led them to reject most or all of the science knowledge presented to them. These results are in line with several recent studies in which conspiracy theorists are not less susceptible to social influence in general [@pummererConspiracyBeliefsMajority; @altayConspiracyBelieversClaim2023].

Our findings are also relevant for the knowledge–attitudes model of trust in science [see, e.g., @bauerWhatCanWe2007]. The fact that even people with low science knowledge trust most of basic science might be taken as showing the power of even a modicum of science knowledge; on the other hand, it also means that attempts at increasing science knowledge in general might not have much effect on trust in science, by contrast with targeting the specific beliefs people are motivated to reject.

Finally, some of the present results also speak to the alienation model [@gauchatCulturalAuthorityScience2011], and more specifically to the need for epistemic autonomy [@frickerTestimonyEpistemicAutonomy2006]. Declaring not trusting science, or endorsing conspiracy theories [@harrisConspiracyTheoriesPopulism2023] might reflect a desire to maintain epistemic autonomy and not appear to ‘blindly’ accept epistemic authority. The acceptance of basic science facts would be reconciled with this need to appear epistemically autonomous by the suggestion that the acceptance stems from independent evaluation instead of trust (however implausible that might be: it’s not clear how participants could independently verify the ratio of glial cells to neurons, say, and few participants appeared to have engaged in even simple forms of verification).

In applied terms, the present results suggest the existence of a vast reservoir of trust in basic science that can be tapped into in science communication. Since people appear so resistant to rejecting basic science, stressing the basic science components of publicly controversial fields, from GMOs to climate change, might help reduce the ‘consensus gaps’ observed in these domains [on climate change, see, e.g. @ranneyClimateChangeConceptual2016].

The present studies have a number of limitations, in particular the lack of representative samples, and the focus on a single country. We hope that future studies will extend our findings to other populations.

\FloatBarrier

# References

::: {#refs}
:::

\newpage

# (APPENDIX) Appendix {.unnumbered}

```{r child = "appendix_acceptance_knowledge.Rmd"}
```

\clearpage

```{r child = "appendix_trust.Rmd"}
```

\clearpage

```{r child = "appendix_conspiracy_thinking.Rmd"}
```

\clearpage

```{r child = "appendix_conspiracy_belief.Rmd"}
```

\clearpage

```{r child = "appendix_knowledge.Rmd"}
```

\clearpage

```{r child = "appendix_by-question.Rmd"}
```

\clearpage

```{r child = "appendix_exp1.Rmd"}
```

\clearpage

```{r child = "appendix_exp2.Rmd"}
```

\clearpage

```{r child = "appendix_exp3.Rmd"}
```

\clearpage

```{r child = "appendix_exp4.Rmd"}
```