### 2025-09-05 README.qmd
add README file and a file with producing additional figures

---
format: gfm
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  fig.retina = 3
)

yaml_metadata <- rmarkdown::yaml_front_matter(here::here("preprint.Rmd"))
title <- yaml_metadata$title
abstract <- yaml_metadata$abstract
```

# `r title`

---

## Abstract

`r abstract`

---

## How to download and replicate

To reproduce the findings and re-run the analysis, do the following:

1. Download and install this repository. You can use GitHub to clone or fork the repository (see the green "Clone or download" button at the top of the GitHub page).
2. Open `everyone_trusts_science_abit.Rproj` to open an [RStudio Project](https://r4ds.had.co.nz/workflow-projects.html).
3. Within the RStudio Project, open the `paper.Rmd` (alternatively, `preprint.Rmd` for a version with authors and affiliations) file and run it.

Please find the [session info below](#session-info).

*Estimated installation time* : If you have R and Rstudio installed, then adding additional packages required in our R scripts is quick (a couple of minutes)

*Estimated run time*: Rendering the `paper.Rmd` document takes less than 1min on our machine (see [session info](#session-info))

## Files glossary {#glossary}

- the two main files are the reproducible source scripts of the paper: `paper.Rmd` is an anonymized version for peer review, `preprint.Rmd` is the version for the public preprint on OSF. 
- all files with the prefix `paper_*` or `preprint_*` are output files that are generated by these source files. 
- `appendix*` files contain the appendices that are automatically added to `manuscript.Rmd`/`preprint.Rmd` when rendering.
- `experiment_*` folders contain all files related to the respective experiment. The `/data` folders contain the raw data (`qualtrics.csv`) and cleaned data (`cleaned.csv`). The cleaning can be reproduced by running the `cleaning.Rmd` files in the folder of each respective experiment. 
- `qualtrics` folders within each experiment folder contains copies of the surveys run via Qualtrics
- `functions/` folder contains functions that we rely upon in our analysis scripts. To keep the `paper.Rmd`/`preprint.Rmd` document neat and readable, we wrote all lengthy code bits as external functions.
- running the `paper.Rmd`/`preprint.Rmd` file will generate an `output` folder which lists all figures generated by the `paper.Rmd` file as `.png` and `.pdf` files.
- `references.bib` contains the references

## Session info {#session-info}

```{r, echo=FALSE}
sessionInfo()
```

## Licenses

**Text and figures:** All prose and images are licensed under Creative Commons ([CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)).

**Code:** All code is licensed under the [MIT License](LICENSE.md).