### 2025-09-05 figures/additional_figures.qmd

---
title: "Additional figures"
format: html
---

```{r}
#| label: setup

library(tidyverse)
library(gt)
library(kableExtra)
library(readxl)
library(wesanderson) # colours
library(gghalves)
library(broom)
library(broom.mixed)
library(ggpubr)
```

```{r functions}
# load plot theme
source("../functions/plot_theme.R")

# load other functions
source("../functions/own_functions.R")
```

```{r}
# read data
exp1_long <- read_csv("../exp_1/data/cleaned_long.csv")
exp2_long <- read_csv("../exp_2/data/cleaned_long.csv")
exp3_long <- read_csv("../exp_3/data/cleaned_long.csv")
exp4_long <- read_csv("../exp_4/data/cleaned_long.csv")

## combine data of all three studies
combined_data <- bind_rows(exp1_long %>% 
                             mutate(study = "1"), 
                           exp2_long %>% 
                             mutate(study = "2"), 
                           exp3_long %>% 
                             mutate(study = "3"),
                           exp4_long %>% 
                             mutate(study = "4")
                           ) %>% 
  ## make a unique id variable
  mutate(id = paste0("study", study, "_", id))
```

```{r}
# make plot data for acceptance
plot_data_acceptance <- combined_data |> 
  drop_na(wgm_sciencegeneral, acceptance) |> 
  # make a labels version of trust in science
  mutate(wgm_sciencegeneral = case_when(
    wgm_sciencegeneral == 1 ~ "1 (Not at all)",
    wgm_sciencegeneral == 2 ~ "2 (Not much)",
    wgm_sciencegeneral == 3 ~ "3 (Some)",
    wgm_sciencegeneral == 4 ~ "4 (A lot)",
    TRUE ~ as.character(wgm_sciencegeneral)  # Handle any other cases (optional)
  )) |> 
  group_by(wgm_sciencegeneral, acceptance) |> 
  summarize(n = n()) |> 
  group_by(wgm_sciencegeneral) |> 
  mutate (share = n/sum(n)) |> 
  mutate_if(is.numeric, round, digits = 3) |> 
  filter(acceptance == "Yes") |> 
  select(-acceptance) |> 
  mutate(outcome = "Acceptance")

# make plot data for knowledge
plot_data_knowledge <- combined_data |> 
  drop_na(wgm_sciencegeneral, knowledge) |> 
  # make a labels version of trust in science
  mutate(wgm_sciencegeneral = case_when(
    wgm_sciencegeneral == 1 ~ "1 (Not at all)",
    wgm_sciencegeneral == 2 ~ "2 (Not much)",
    wgm_sciencegeneral == 3 ~ "3 (Some)",
    wgm_sciencegeneral == 4 ~ "4 (A lot)",
    TRUE ~ as.character(wgm_sciencegeneral)  # Handle any other cases (optional)
  )) |> 
  group_by(wgm_sciencegeneral, knowledge) |> 
  summarize(n = n()) |> 
  group_by(wgm_sciencegeneral) |> 
  mutate (share = n/sum(n)) |> 
  mutate_if(is.numeric, round, digits = 3) |> 
  filter(knowledge == TRUE) |> 
  select(-knowledge) |> 
  mutate(outcome = "Knowledge")

# combine all data
plot_data <- bind_rows(plot_data_acceptance, plot_data_knowledge) |> 
  mutate(outcome = factor(outcome, levels = c("Knowledge", "Acceptance")
                          )
         )

absolute_numbers <- combined_data %>% 
  group_by(wgm_sciencegeneral) %>% 
  summarise(n = n(), 
            n_participants = n_distinct(id))
```

```{r}
ggplot(plot_data |> 
         mutate(outcome = factor(ifelse(outcome == "Acceptance", "Acceptance", "Knowledge"),
                                 levels = c("Knowledge", "Acceptance"))), 
       aes(x = wgm_sciencegeneral, y = share, fill = outcome)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = scales::percent_format(), breaks = seq(0, 1, 0.1), limits = c(0, 1.1)) +
  geom_text(aes(label = paste0(round(share * 100, 1), "%")),
            position = position_dodge(width = 0.9),
            vjust = -0.5, size = 3) +
  scale_fill_viridis_d(option = "cividis") +
  labs(x = "Trust in science", fill = "Share of", y = "") +
  plot_theme +
  geom_bracket(data = absolute_numbers,
               aes(xmin = as.numeric(factor(wgm_sciencegeneral)) - 0.4, 
                   xmax = as.numeric(factor(wgm_sciencegeneral)) + 0.4, 
                   y.position = 1.05, label =  paste0("n = ", n_participants)),
               inherit.aes = FALSE,
               step.increase = 0) +
  theme(legend.position = "top")
```